### 2023-10-11 paper1-INTERACT-analysis.Rmd
Trying to create a public repo for the INTERACT analysis

---
title: "paper1-INTERACT-analysis"
author: "Shelby Sturrock"
date: "September 10, 2023"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE,results=FALSE,echo=FALSE}
  knitr::opts_chunk$set(echo = TRUE)
  # install.packages("table1")
  # install.packages("effects")
  # install.packages("stargazer")
  library(dplyr)
  library(lubridate)
  library(ggplot2)
  library(tableone)
  library(table1)
  library(tidyr)
  library(lme4)
  library(gridExtra)
  library(MASS)
  library(sjPlot)
  library(effects)
  library(stargazer)
  library(car)
  knitr::opts_knit$set(root.dir = '/Users/shelbysturrock/Library/Mobile Documents/com~apple~CloudDocs/Documents/Documents - Shelby’s MacBook Pro/School/UofT/Dissertation/PAPER 1 - INTERACT')
  knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Paper 1: Analysis

## 1. Prepare data

### Sensdoc data

Read in SD TOP: 
```{r,results="hide"}
  sd_pa<-read.csv("sensdoc-wearing-2023.csv") # 5575
```

### Set levels for categorical variables (and re-categorize per MW comments)

Format date as date
```{r}
  sd_pa$date<-as.Date(sd_pa$date)
```

Age category (other paper used 18-24, 25-44, 45 to 64, and 65+)
```{r,echo=FALSE}
  sd_pa$age_cat<-factor(sd_pa$age_cat,
                  levels=c("18-29","30-49","50-69","70+"))
  table(sd_pa$age_cat)
```

Income. Combine <50,000 into one category and re-order the levels 
```{r,echo=FALSE}
  sd_pa$income<-factor(sd_pa$income,
                levels=c("<$50,000","$50,000 to $99,999",
                          "$100,000 or more","don't know/prefer not to answer"))
  summary(as.factor(sd_pa$income))
```

Gender:  
```{r,echo=FALSE}
  sd_pa$gender<-factor(sd_pa$gender,
                       levels=c("man","woman","other"))
  summary(as.factor(sd_pa$gender))
```

Children: 
```{r,echo=FALSE}
  sd_pa$children<-factor(sd_pa$children,levels=c("yes","no"))
  summary(as.factor(sd_pa$children))
```

Employment status
```{r,echo=FALSE}
  sd_pa$employment<-factor(sd_pa$employment,
                           levels=c("employed","retired and not working","unemployed","other"))
  summary(as.factor(sd_pa$employment))
```

Education: 
```{r,echo=FALSE}
  sd_pa$education<-factor(sd_pa$education,
                          levels=c("university or graduate degree","secondary school",
                                   "trade school or college","don't know/prefer not to answer"))
  summary(as.factor(sd_pa$education))
```

Create categorical stringency index, for assessing balance across levels of stringency 
```{r}
  summary(sd_pa[sd_pa$wave_id=="wave 2",]$stringencyIndex)
    # min = 25.90, 1st qu. = 39.40, median = 47.00, 3rd qu. = 50.70, max = 66.80

# create a categorical version: <40, 40-50, >50
  sd_pa$si_cat<-as.factor(ifelse(sd_pa$stringencyIndex <= 41, "0-41",
                          ifelse(sd_pa$stringencyIndex > 41 & sd_pa$stringencyIndex <=47, "42-47",
                          ifelse(sd_pa$stringencyIndex > 47, "48-100",NA))))
  summary(sd_pa[sd_pa$wave_id=="wave 2",]$si_cat)
```

### Data cleaning: Creating sub-samples

*sd_ppl: Person-level dataset*: used in version 1 of Table 1 (with person-level characteristics)
```{r}
  sd_ppl<-sd_pa %>% ungroup() %>% 
                    dplyr::select(interact_id,income,gender,age,age_cat,children,
                           education,employment,wave_id,city_id,both)
  sd_ppl<-distinct(sd_ppl)
  table(sd_ppl$wave_id,sd_ppl$city_id) # one participant from Saskatoon in wave 1 has gone missing?
```

*sd_valid: Participant-days of data with >= 600 minutes of wear time*: this is used in the main analyses and table 1
```{r}
  sd_valid<-sd_pa[sd_pa$wear>=600,]
  length(sd_pa[sd_pa$wear<600,]$interact_id)
```

*sd_valid_ppl: People with at least one valid participant-day (>= 600 minutes of wear time)*
```{r}
  sd_ppl_valid<-sd_valid %>% ungroup() %>% 
                    dplyr::select(interact_id,income,gender,age,age_cat,children,
                           education,employment,wave_id,city_id,both)
  sd_ppl_valid<-distinct(sd_ppl_valid)
  table(sd_ppl_valid$wave_id) # one participant from Saskatoon in wave 1 has gone missing?
```

*sd_valid_both: valid person-days from participants that contributed at least one day of VALID data in both waves* this is used for sensitivity analysis I: filtered to valid person-days (sd_valid) and then filtered to participants with data in both waves (sd_valid_both) 
```{r}
  valid_both_waves<-sd_valid %>% dplyr::select(interact_id,wave_id) %>% distinct() %>% 
                                        group_by(interact_id) %>%
                                        mutate(count=length(unique(wave_id))) %>% filter(count>1)
  sd_valid_both<-sd_valid[sd_valid$interact_id %in% valid_both_waves[valid_both_waves$count==2,]$interact_id ,]
  length(unique(sd_valid_both$interact_id))
```

### Stringency index

Read in data
```{r}
  si<-read.csv("./Data/IRPP stringency index/2021 09 22 INDEX Centre of Excellence COVID Policy Data_New Codebook.csv")
  si$date<-as.Date(si$date)
  si$province<-si$Province.Territory
```

Rescale stringencyIndex by dividing values by 10
```{r}
  si$stringencyIndex<-si$stringencyIndex/10
  sd_pa$stringencyIndex<-sd_pa$stringencyIndex/10
```

Rescale wear time
```{r}
  sd_pa$wearRS<-sd_pa$wear/100
  sd_valid$wearRS<-sd_valid$wear/100
```

Plot of stringency index over time (in black and white)
```{r,echo=FALSE}
  si_plot<-ggplot(si[si$province %in% c("British Columbia","Saskatchewan","Quebec") & 
                      si$date <= "2021-02-28",],aes(x=date,y=stringencyIndex*10,color=province)) +
           geom_rect(aes(xmin=ymd('2020-09-10'),
                  xmax = ymd('2021-02-28'),
                  ymin = -Inf,
                  ymax = Inf), fill = 'grey95', color='grey95',alpha = 0.75) +
           geom_line(size=0.6) + 
           theme_classic() + 
           scale_colour_grey(labels=c("British Columbia\n(Vancouver)\n",
                                      "Quebec\n(Montreal)\n",
                                      "Saskatchewan\n(Saskatoon)\n")) +
           labs(#title="Stringency of COVID-19 restrictions over time",
                 x="Date",y="Stringency index") +
           guides(color=guide_legend(title="Province (City)\n")) +
           theme(legend.title=element_text(face="bold",size=9),
                 legend.text = element_text(size=8),
                 axis.title = element_text(face="bold",size=9),
                 axis.text = element_text(size=8),
                 plot.title = element_text(face="bold",size=10,hjust=0.5)) +
           annotate(geom="text", x=as.Date('2020-12-04'), y=10, label="Wave 2 data collection period",
                    color="grey30",size=2.75)

  si_plot

  ggsave(si_plot,file="figures//figure4.pdf",width=5.5,height=3.25)
  ggsave(si_plot,file="submission to IJE//figure4.pdf",width=175,height=95,unit="mm") # for IJBNPA
```

## 2. Assess for balance across exposures

### Create Table 1:

*Columns 1 and 2: participants who contributed at least one valid day of data, by wave (sd_ppl_valid)*
```{r, echo=FALSE}
  balance_ppl_by_wave_full<-CreateTableOne(vars=c(
                                        "city_id","gender","age_cat","income",
                                        "children","education","employment"),
                                 strata="wave_id",
                                 data=sd_ppl_valid,test=FALSE,
                                 includeNA=TRUE)
  kableone(balance_ppl_by_wave_full)
  #print(kableone(balance_ppl_by_wave_full,quote=TRUE,noSpaces=TRUE))
```

*Columns 3 and 4: valid person-days, by wave (sd_valid)*
Wear time and city are very different between waves. Some differences in age (older in wave 2), some differences in income (fewer young people, more older people in wave 2) (vs. sd_pa: all person-days, by wave)
filtering out invalid days introduces an imbalance in education levels
different: wear, city, age, income, employment status, education 
```{r, echo=FALSE}
  balance_days_by_wave_valid<-CreateTableOne(vars=c(
                                        "wear","mean_mvpa_sd","mean_temp","total_precip",
                                        "city_id","gender","age_cat","income",
                                        "children","education","employment"),
                                 strata="wave_id",
                                 data=sd_valid,test=FALSE,
                                 includeNA=TRUE)
   kableone(balance_days_by_wave_valid)
   #print(kableone(balance_days_by_wave_valid,quote=TRUE,noSpaces=TRUE))
```

*Columns 5 and 6: valid participant-days from wave 2, stratified by stringency index (sd_valid)*: 
CLEAR imbalances of gender, age, kids, education and employment status (not income)
```{r, echo=FALSE}
  balance_by_SI_full<-CreateTableOne(vars=c(#"mean_wave_mvpa","mean_wave_wear",
                                        "wear","mean_mvpa_sd","mean_temp","total_precip",
                                        "city_id","gender","age_cat","income",
                                        "children","education","employment"),
                                 strata="si_cat",
                                 data=sd_valid,test=FALSE,
                                 includeNA=TRUE)
  kableone(balance_by_SI_full)
  #print(kableone(balance_by_SI_full,quote=TRUE,noSpaces=TRUE))
```

### Create Supplementary Table 1:

*Columns 1 and 2: Participants with at least one valid day of data in BOTH waves (objective 1, sensitivity analysis I; sd_valid_both*: 
imbalances are slightly different here, too. In theory, this group should be the same in both waves, but this is a person-day dataset and there are people who contributed different number of (valid) days in each wave
different: wear, gender, income, 
same: city, education, employment, kids 
```{r, echo=FALSE}
  balance_days_by_wave_valid_both<-CreateTableOne(vars=c(
                                        "wear","mean_mvpa_sd","mean_temp","total_precip",
                                        "city_id","gender","age_cat","income",
                                        "children","education","employment"),
                                 strata="wave_id",
                                 data=sd_valid_both,test=FALSE)
  kableone(balance_days_by_wave_valid_both)
  #print(kableone(balance_days_by_wave_valid_both,quote=TRUE,noSpaces=TRUE))
```

*Columns 3 and 4: Valid person-days collected between October 3 and February 20 (objective 1, sensitivity analysis II; sd_valid)*
```{r, echo=FALSE}
  balance_days_by_wave_septFeb<-CreateTableOne(vars=c(
                                        "wear","mean_mvpa_sd","mean_temp","total_precip",
                                        "city_id","gender","age_cat","income",
                                        "children","education","employment"),
                                 strata="wave_id",
                                 data=sd_valid[
                                   (sd_valid$wave_id == "wave 1" & sd_valid$date >= "2018-10-03") |
                                   (sd_valid$wave_id == "wave 2" & sd_valid$date <= "2021-02-20"),],test=FALSE)
  kableone(balance_days_by_wave_septFeb)
  #print(kableone(balance_days_by_wave_septFeb,quote=TRUE,noSpaces=TRUE))
```

Rescale temperature and precipitation
```{r}
  sd_pa$total_precip<-sd_pa$total_precip/10
  sd_pa$mean_temp<-sd_pa$mean_temp/10
```

## 3. OBJECTIVE 1: Impact of the pandemic on daily MVPA

### Descriptive figures

**Figure 1**: Distribution of daily minutes of wear time and moderate-to-vigorous physical activity among valid person-days, by wave
```{r,echo=FALSE}
wave.labs<-c("Wave 1","Wave 2")
names(wave.labs)<-c("wave 1","wave 2")

  wear_plot<-ggplot(sd_valid,aes(x=wear)) + 
    geom_histogram(bins=50,aes(fill=wave_id,color=wave_id)) +
    theme_classic() +
    labs(x="Minutes",
         y="Frequency",
         title="Wear time" #,
         #subtitle="Wave 1 (dark blue) vs. Wave 2 (light blue)"
         ) +
    geom_vline(xintercept=600,color="black",linetype="dashed") +
    facet_grid(vars(wave_id),
               labeller = labeller(wave_id=wave.labs)) +
    # scale_fill_manual(values=c("wave 1"="#0a9ffe", # green = "#81DA66"
    #                             "wave 2"="#fe7701"), # darker orange #fe7701 #ff8c09
    #                    labels=c("Wave 1","Wave 2")) + 
    scale_fill_grey(labels=c("Wave 1","Wave 2")) +
    scale_color_grey(labels=c("Wave 1","Wave 2"))+
    theme(axis.title.x = element_text(face="bold",
                                       size=10),
          axis.text.x = element_text(angle=90,vjust = 0.5, hjust=1,size=8),
        # axis.title.y = element_text(#face="bold",
        #                              size=10),
          axis.title.y = element_blank(),
          axis.text.y = element_text(size=8),
          plot.title = element_text(face=2,
                                    hjust=0.5,
                                    size=10),
          legend.position="none",
          strip.text.y = element_text(face=2))

  mvpa_plot<-ggplot(sd_valid,aes(x=mean_mvpa_sd)) + 
    geom_histogram(bins=50,aes(fill=wave_id,color=wave_id)) +
    theme_classic() +
    labs(x="Minutes",
         y="Frequency",
         title="Moderate-to-vigorous physical activity" #,
         #subtitle="Wave 1 (dark blue) vs. Wave 2 (light blue)"
         ) +
    facet_grid(vars(wave_id),
               labeller = labeller(wave_id=wave.labs)) +
    # scale_fill_manual(values=c("wave 1"="#0a9ffe", # green = "#81DA66"
    #                             "wave 2"="#fe7701"), # darker orange #fe7701 #ff8c09
    #                    labels=c("Wave 1","Wave 2")) + 
    scale_fill_grey(labels=c("Wave 1","Wave 2")) +
    scale_color_grey(labels=c("Wave 1","Wave 2"))+
    theme(axis.title.x = element_text(face="bold",
                                       size=10),
          axis.text.x = element_text(angle=90,vjust = 0.5, hjust=1,size=8),
        # axis.title.y = element_text(#face="bold",
        #                              size=10),
          axis.title.y = element_blank(),
          axis.text.y = element_text(size=8),
          plot.title = element_text(face=2,
                                    hjust=0.5,
                                    size=10),
          legend.position="none",
          strip.text.y = element_text(face=2))

  # titleMain=ggpubr::text_grob("Distribution of daily minutes of wear time and moderate-to-vigorous physical activity:\nWave 1 (2018-2019) vs. Wave 2 (2020-2021)\n",
  #                             size =10,face="bold")  
  titleY=ggpubr::text_grob("Frequency",
                            size =10,rot=90,face="bold")

  figure1<-grid.arrange(wear_plot,mvpa_plot,ncol=2,
               #top=titleMain, # removing title for IJBNPA (titles should be in the text, not in the figure)
               left=titleY)

  ggsave(figure1,file="figures//figure1.pdf",width=6,height=4)
  ggsave(figure1,file="submission to IJE//figure1.pdf",width=175,height=100,unit="mm")
```

**Figure 2:** Trend in daily minutes of moderate-to-vigorous physical activity for each wave
```{r,echo=FALSE,warning=FALSE}
  sd_valid$doy3<-sd_valid$date
  sd_valid$doy3<-gsub("2018|2020","2000",sd_valid$doy3)
  sd_valid$doy3<-gsub("2019|2021","2001",sd_valid$doy3)
  sd_valid$doy3<-as.Date(sd_valid$doy3)

  figure2<-ggplot(sd_valid[sd_valid$wave_id=="wave 1",],aes(x=doy3,y=mean_mvpa_sd,color=wave_id)) +
                          geom_point(alpha=0.15,size=1)+
                          geom_point(data=sd_valid[sd_valid$wave_id=="wave 2",],aes(x=doy3,y=mean_mvpa_sd),
                                     alpha=0.15,size=1) +
                          scale_colour_manual(values=c("wave 1"="grey65",
                                                       "wave 2"="black"),
                                              labels=c("Wave 1 (2018-2019)","Wave 2 (2020-2021)")) +
                          geom_smooth(method=lm)+
                          geom_smooth(data=sd_valid[sd_valid$wave_id=="wave 2",],aes(x=doy3,y=mean_mvpa_sd),method=lm) +
                          theme_classic()+
                          labs(#title = "Daily minutes of moderate-to-vigorous physical activity by day of the year: ", # removing title for IJBNPA
                               #subtitle = "Wave 1 (2018-2019) vs. Wave 2 (2020-2021)",
                               y = "Daily minutes of moderate-to-vigorous physical activity",
                               x = "Date")+
                          guides(color=guide_legend("")) +
                          theme(axis.title = element_text(face="bold",size=10),
                                axis.text.x = element_text(angle=90,vjust = 0.5, hjust=1,size=8),
                                axis.text.y = element_text(size=8),
                                plot.title = element_text(face="bold",hjust=0.5,size=10),
                                plot.subtitle = element_text(hjust=0.5,size=10),
                                legend.position="bottom",
                                legend.key.size = unit(1,"line"),
                                legend.spacing.x = unit(0.25, 'cm'),
                                legend.margin=margin(-3, 0, 0, 0),
                                legend.text = element_text(size=8,margin = margin(r = 0.4, unit = 'cm'))) 
  figure2
  ggsave(figure2,file="figure2.pdf",width=6,height=4)
  ggsave(figure2,file="submission to IJE//figure2.pdf",width=175,height=120,unit="mm")
```

### I. Main effect models (results presented in Table 2)

#### Unadjusted model (Table 2, Row 1)

*Model*
```{r}
  wave_abs_valid_unadj<-glmer.nb(mean_mvpa_sd~ wave_id + (1|interact_id),
                                 data=sd_valid) 
  #summary(wave_abs_valid_unadj)
```

*Results*
```{r,echo=FALSE}
# effect estimate:
  coef4<-as.data.frame(summary(wave_abs_valid_unadj)$coefficients)
  print(paste("effect estimate: ",exp(coef4[2,]$Estimate),sep=""))

# confidence interval: 
  ci4<-exp(confint(wave_abs_valid_unadj, c("wave_idwave 2"), level = 0.95))
  print("confidence interval: ")
  print(ci4)

# number of observations for wave 1 and wave 2?
  # colSums(model.matrix(wave_abs_valid_unadj)) # total = 3260, wave 1 = 3010, wave 2 = 250
```

#### Adjusted for unbalanced factors (Table 2, Row 2)

*Assess for balance betwween waves to determine that to adjust for in the adjusted model*
```{r,echo=FALSE}
  kableone(balance_days_by_wave_valid)
```

*Model: Adjusting for city, age, employment status, income, temp, precipitation and wear time*
```{r,warning=FALSE}
  wave_covar_valid_fully<-glmer.nb(mean_mvpa_sd~ wave_id + city_id + wearRS +
                                                 age_cat + employment + income +
                                                 mean_temp + total_precip + 
                                                 (1|interact_id),
                          data=sd_valid[sd_valid$age_cat!="missing",])
```

*Results*
```{r,echo=FALSE}
# effect estimate:
  coef7<-as.data.frame(summary(wave_covar_valid_fully)$coefficients)
  print(paste("effect estimate: ",exp(coef7[2,]$Estimate),sep=""))

# confidence interval: 
  ci7<-exp(confint(wave_covar_valid_fully, c("wave_idwave 2"), level = 0.95))
  print("confidence interval: ")
  print(ci7)

# number of observations for wave 1 and wave 2?
  # colSums(model.matrix(wave_covar_valid_fully)) # total = 3142 (3260-118), wave 1 = 2893, wave 2 = 249
```

### II. Sensitivity analyses (Supplementary Table 2)

#### A. Sensitivity analysis 1 (Supplementary Table 2, Row 2)

Participants-days from participants that contributed data in both waves with 600+ minutes of wear time, adjusted for unbalanced factors (gender, age, household income, mean daily temperature, daily precipitation, daily wear time)

**Model**
```{r,warning=FALSE}
  both_valid_adj_wear<-glmer.nb(mean_mvpa_sd~ wave_id + 
                                              gender + age_cat + income +
                                              mean_temp + total_precip + wear +
                                       (1|interact_id),
                          data=sd_valid_both) 
```

**Results**
```{r,echo=FALSE}
# effect estimate: 
  coefSens1<-as.data.frame(summary(both_valid_adj_wear)$coefficients)[2,]
  print(paste("effect estimate: ",round(exp(coefSens1$Estimate),digits=2),sep=""))

# confidence interval: 
  ciSens1<-exp(confint(both_valid_adj_wear, c("wave_idwave 2"), level = 0.95))
  print("confidence interval: ")
  print(round(ciSens1,digits=2))

# number of observations for wave 1 and wave 2?
  colSums(model.matrix(both_valid_adj_wear)) # total = 454 (459-5), wave 1 = 307, wave 2 = 147
```

#### B. Sensitivity analysis 2 (Supplementary Table 2, Row 3)

Filter data from wave 1 and wave 2 to the same date range

**Dates of data collection for wave 1 vs. wave 2**
```{r,echo=FALSE}
  paste("dates for wave 1 data collection: ",min(sd_valid[sd_valid$wave_id=="wave 1",]$date)," to ",
                                             max(sd_valid[sd_valid$wave_id=="wave 1",]$date),sep="")
  paste("dates for wave 2 data collection: ",min(sd_valid[sd_valid$wave_id=="wave 2",]$date)," to ",
                                             max(sd_valid[sd_valid$wave_id=="wave 2",]$date),sep="")
```

**Create new dataframe restricted to valid days between October 3 and February 20 (the range of dates on which person-days with 600+ minutes of wear time begin)**
```{r}
  sens_valid<-sd_valid[(sd_valid$wave_id == "wave 1" & sd_valid$date >= "2018-10-03") |
                       (sd_valid$wave_id == "wave 2" & sd_valid$date <= "2021-02-20"),]
```

**Double check new data frame: Note the time frame matches for both wave 1 and wave 2**
```{r,echo=FALSE}
  paste("dates for wave 1 data collection: ",min(sens_valid[sens_valid$wave_id=="wave 1",]$date)," to ",
                                             max(sens_valid[sens_valid$wave_id=="wave 1",]$date),sep="")
  paste("dates for wave 2 data collection: ",min(sens_valid[sens_valid$wave_id=="wave 2",]$date)," to ",
                                             max(sens_valid[sens_valid$wave_id=="wave 2",]$date),sep="")
```

**Model**
```{r,warning=FALSE}
  wave_covar_valid_fully_sens<-glmer.nb(mean_mvpa_sd~ wave_id +
                                                      gender + age_cat + employment + income + education +
                                                      city_id + wear +
                                       (1|interact_id),
                          data=sens_valid[sens_valid$age_cat!="missing",])
  summary(wave_covar_valid_fully_sens)
```

**Results**
```{r,echo=FALSE}
# effect estimate:
  coefSens2<-as.data.frame(summary(wave_covar_valid_fully_sens)$coefficients)[2,]
  print(paste("effect estimate: ",round(exp(coefSens2$Estimate),digits=2),sep=""))

# confidence interval: 
  ciSens2<-exp(confint(wave_covar_valid_fully_sens, c("wave_idwave 2"), level = 0.95))
  print("confidence interval: ")
  print(round(ciSens2,digits=2))
```

### III. Effect modification

#### Assess for effect modification (Supplementary Table 3)

I'll run stratified models to determine if the effect size or direction of the exposure variable (wave_id) varies across levels of each independent variable. These models will be run on valid person-days (600+ minutes of wear time) from both waves.

**Overall, for reference (row 1 supplementary table 2)**
```{r,warning=FALSE,echo=FALSE}
  overall_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid) 
# Comparison: 
  print(paste("overall: ",round(exp(as.data.frame(summary(overall_valid_adj)$coefficients)$Estimate[2]),digits=2)),sep="")
```

**Gender**
```{r,echo=FALSE,warning=FALSE}
# models (men, women):
  gender1_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$gender=="man",]) 
  gender2_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$gender=="woman",])

# effect estimates:
  gender1_coef<-round(exp(as.data.frame(summary(gender1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  gender2_coef<-round(exp(as.data.frame(summary(gender2_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  gender1_CI<-round(exp(confint(gender1_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  gender2_CI<-round(exp(confint(gender2_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
```

Distribution of person-days by wave and gender
```{r,echo=FALSE}
  addmargins(table(as.factor(sd_valid$gender),as.factor(sd_valid$wave_id)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
  print(paste("men: ",gender1_coef[1]," (",gender1_CI[1],", ",gender1_CI[2],")",sep=""))
  print(paste("women: ",gender2_coef[1]," (",gender2_CI[1],", ",gender2_CI[2],")",sep=""))
```

**Age**
```{r,echo=FALSE,warning=FALSE}
# models (18-29, 30-49, 50-69, 70+) 
  age1_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$age_cat=="18-29",]) 
  age2_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$age_cat=="30-49",]) 
  age3_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$age_cat=="50-69",]) 
  age4_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$age_cat=="70+",])

# effect estimates:
  age1_coef<-round(exp(as.data.frame(summary(age1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  age2_coef<-round(exp(as.data.frame(summary(age2_valid_adj)$coefficients)$Estimate[2]),digits=2)
  age3_coef<-round(exp(as.data.frame(summary(age3_valid_adj)$coefficients)$Estimate[2]),digits=2)
  age4_coef<-round(exp(as.data.frame(summary(age4_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  age1_CI<-round(exp(confint(age1_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  age2_CI<-round(exp(confint(age2_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  age3_CI<-round(exp(confint(age3_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  age4_CI<-round(exp(confint(age4_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
```

Distribution of person-days by wave and age
```{r,echo=FALSE}
  addmargins(table(as.factor(sd_valid$age_cat),as.factor(sd_valid$wave_id)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
  print(paste("18-29: ",age1_coef[1]," (",age1_CI[1],", ",age1_CI[2],")",sep=""))
  print(paste("30-49: ",age2_coef[1]," (",age2_CI[1],", ",age2_CI[2],")",sep=""))
  print(paste("50-69: ",age3_coef[1]," (",age3_CI[1],", ",age3_CI[2],")",sep=""))
  print(paste("70+: ",   age4_coef[1]," (",age4_CI[1],", ",age4_CI[2],")",sep=""))
```

**Income**
```{r,echo=FALSE,warning=FALSE}
# models (<50,000, 50,000-99,999, 100,000-149,999, 150,000+)
  income1_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$income=="<$50,000",]) 
  income2_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$income=="$50,000 to $99,999",]) 
  income3_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$income=="$100,000 or more",]) 
  income4_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$income=="don't know/prefer not to answer",])

# effect estimates:
  income1_coef<-round(exp(as.data.frame(summary(income1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  income2_coef<-round(exp(as.data.frame(summary(income2_valid_adj)$coefficients)$Estimate[2]),digits=2)
  income3_coef<-round(exp(as.data.frame(summary(income3_valid_adj)$coefficients)$Estimate[2]),digits=2)
  income4_coef<-round(exp(as.data.frame(summary(income4_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  income1_CI<-round(exp(confint(income1_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  income2_CI<-round(exp(confint(income2_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  income3_CI<-round(exp(confint(income3_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  income4_CI<-round(exp(confint(income4_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
```

Distribution of person-days by wave and income
```{r,echo=FALSE}
  addmargins(table(as.factor(sd_valid$income),as.factor(sd_valid$wave_id)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
  print(paste("<$50,000: ",income1_coef[1]," (",income1_CI[1],", ",income1_CI[2],")",sep=""))
  print(paste("$50,000 to $99,999: ",income2_coef[1]," (",income2_CI[1],", ",income2_CI[2],")",sep=""))
  print(paste("$100,000 or more: ",income3_coef[1]," (",income3_CI[1],", ",income3_CI[2],")",sep=""))
  print(paste("don't know/prefer not to answer: ",income4_coef[1]," (",income4_CI[1],", ",income4_CI[2],")",sep=""))
```

**Education**
```{r,echo=FALSE,warning=FALSE}
# models (university or graduate degree, secondary school, trade school or college)
  education1_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$education=="university or graduate degree",])
  education2_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$education=="secondary school",]) 
  education3_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$education=="trade school or college",])

# effect estimates:
  education1_coef<-round(exp(as.data.frame(summary(education1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  education2_coef<-round(exp(as.data.frame(summary(education2_valid_adj)$coefficients)$Estimate[2]),digits=2)
  education3_coef<-round(exp(as.data.frame(summary(education3_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  education1_CI<-round(exp(confint(education1_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  education2_CI<-round(exp(confint(education2_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  education3_CI<-round(exp(confint(education3_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
```

Distribution of person-days by wave and education
```{r,echo=FALSE}
  addmargins(table(as.factor(sd_valid$education),as.factor(sd_valid$wave_id)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
  print(paste("university or graduate degree: ",
              education1_coef[1]," (",education1_CI[1],", ",education1_CI[2],")",sep=""))
  print(paste("secondary school: ",
              education2_coef[1]," (",education2_CI[1],", ",education2_CI[2],")",sep=""))
  print(paste("trade school or college: ",
              education3_coef[1]," (",education3_CI[1],", ",education3_CI[2],")",sep=""))
```

**Employment status**
```{r,echo=FALSE,warning=FALSE}
# models (employed, retired and not working, unemployed, other): 
  employment1_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$employment=="employed",]) 
  employment2_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$employment=="retired and not working",]) 
  employment3_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$employment=="unemployed",]) 
  employment4_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$employment=="other",])

# effect estimates:
  employment1_coef<-round(exp(as.data.frame(summary(employment1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  employment2_coef<-round(exp(as.data.frame(summary(employment2_valid_adj)$coefficients)$Estimate[2]),digits=2)
  employment3_coef<-round(exp(as.data.frame(summary(employment3_valid_adj)$coefficients)$Estimate[2]),digits=2)
  employment4_coef<-round(exp(as.data.frame(summary(employment4_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  employment1_CI<-round(exp(confint(employment1_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  employment2_CI<-round(exp(confint(employment2_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  employment3_CI<-round(exp(confint(employment3_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  employment4_CI<-round(exp(confint(employment4_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
```

Distribution of person-days by wave and employment
```{r,echo=FALSE}
  addmargins(table(as.factor(sd_valid$employment),as.factor(sd_valid$wave_id)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
  print(paste("employed: ",
              employment1_coef[1]," (",employment1_CI[1],", ",employment1_CI[2],")",sep=""))
  print(paste("retired and not working: ",
              employment2_coef[1]," (",employment2_CI[1],", ",employment2_CI[2],")",sep=""))
  print(paste("unemployed: ",
              employment3_coef[1]," (",employment3_CI[1],", ",employment3_CI[2],")",sep=""))
  print(paste("other: ",
              employment4_coef[1]," (",employment4_CI[1],", ",employment4_CI[2],")",sep=""))
```

**Has children**
```{r,echo=FALSE,warning=FALSE}
# models:
  children1_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$children=="yes",]) 
  children2_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$children=="no",])

# effect estimates:
  children1_coef<-round(exp(as.data.frame(summary(children1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  children2_coef<-round(exp(as.data.frame(summary(children2_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  children1_CI<-round(exp(confint(children1_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  children2_CI<-round(exp(confint(children2_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
```

Distribution of person-days by wave and employment
```{r,echo=FALSE}
  addmargins(table(as.factor(sd_valid$children),as.factor(sd_valid$wave_id)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
  print(paste("has children at home: ",
              children1_coef[1]," (",children1_CI[1],", ",children1_CI[2],")",sep=""))
  print(paste("does not have children at home:",
              children2_coef[1]," (",children2_CI[1],", ",children2_CI[2],")",sep=""))
```

**City**
```{r,echo=FALSE,warning=FALSE}
# models: 
  city1_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$city_id=="montreal",])
  city2_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$city_id=="saskatoon",]) 
  city3_valid_adj<-glmer.nb(mean_mvpa_sd~ wave_id + wearRS + (1|interact_id),
                              data=sd_valid[sd_valid$city_id=="vancouver",])

# effect estimates:
  city1_coef<-round(exp(as.data.frame(summary(city1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  city2_coef<-round(exp(as.data.frame(summary(city2_valid_adj)$coefficients)$Estimate[2]),digits=2)
  city3_coef<-round(exp(as.data.frame(summary(city3_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  city1_CI<-round(exp(confint(city1_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  city2_CI<-round(exp(confint(city2_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")),digits=2)
  city3_CI<-round(exp(confint(city3_valid_adj, c("wave_idwave 2"), level = 0.95,method="boot")) ,digits=2)
```

Distribution of person-days by wave and employment
```{r,echo=FALSE}
  addmargins(table(as.factor(sd_valid$city_id),as.factor(sd_valid$wave_id)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
  print(paste("montreal: ",
              city1_coef[1]," (",city1_CI[1],", ",city1_CI[2],")",sep=""))
  print(paste("saskatoon: ",
              city2_coef[1]," (",city2_CI[1],", ",city2_CI[2],")",sep=""))
  print(paste("vancouver: ",
              city3_coef[1]," (",city3_CI[1],", ",city3_CI[2],")",sep=""))
```

#### Plot effect modifiers of interest (Figure 3): Income, employment status and education

Plot interaction effect using effects package (https://cran.r-project.org/web/packages/effects/effects.pdf)

Note: these models are run on person-days with 600+ minutes of wear time and thus we'll control for factors that are unbalanced within this sub-sample

**Create income plot**
```{r,echo=FALSE,results=FALSE,warning=FALSE}
# Add a income-wave interaction term to the adjusted model  
income_covar_valid_final<-glmer.nb(mean_mvpa_sd~ wave_id + city_id + wave_id*income + wearRS +
                                                age_cat + employment + income +
                                                mean_temp + total_precip + 
                                                (1|interact_id),
                          data=sd_valid[sd_valid$age_cat!="missing",])

# estimate the effect
income_Inter.HandPick <- effect('wave_id*income', income_covar_valid_final,
                                            xlevels=list(wave_id = c("wave 1","wave 2"),
                                            income = c("<$50,000","$50,000 to $99,999","$100,000 or more",
                                                       "don't know/prefer not to answer"
                                                       )),
                                            se=TRUE, confidence.level=.95, typical=mean)

# put results in data frame 
income_Inter.HandPick <- as.data.frame(income_Inter.HandPick)

# create factor variables for both wave_id and income             
income_Inter.HandPick$Income <- factor(income_Inter.HandPick$income,
                      levels=c("<$50,000","$50,000 to $99,999","$100,000 or more","don't know/prefer not to answer"
                               ),
                      labels=c("<$50,000","$50,000-$99,999","$100,000+","No answer"
                               ))
income_Inter.HandPick$wave <- factor(income_Inter.HandPick$wave_id,
              levels=c("wave 1","wave 2"),
              labels=c("Wave 1", "Wave 2"))

# plot the interaction effect
income_Plot.HandPick<-ggplot(data=income_Inter.HandPick, aes(x=wave, y=fit, group=Income))+
      geom_point(size=1, aes(color=Income),position=position_dodge(0.5))+
      ylim(0,100) +
      geom_errorbar(aes(ymin=lower, ymax=upper,color=Income), width=.3,size=0.6, position=position_dodge(0.5))+
      scale_colour_grey()+
      labs(title="Household income")  +
      guides(color=guide_legend("Income",nrow=4,byrow=TRUE,
                                 keywidth=0.15,
                                 keyheight=0.1,
                                 default.unit="inch"
                                 )) +
                     theme_classic() + theme(plot.title=element_text(size=12,
                                                                     face=2,
                                                                     hjust=0.5),
                                       axis.text.x=element_text(size=10,
                                                              angle = 0,
                                                              face= "bold"),
                                       axis.text.y=element_text(size=10,
                                                              angle = 0),
                                       legend.title=element_blank(),
                                       legend.text = element_text(size=8.5,margin = margin(r = 0.24, unit = 'cm')),
                                       axis.title.x = element_blank(),
                                       axis.title.y = element_blank(),
                                       legend.key.size = unit(0.5,"line"),
                                       legend.position="bottom",
                                       legend.spacing.x = unit(0.06, 'cm'),
                                       legend.margin=margin(-5, 0, 0, 0),
                                       plot.margin = unit(c(12, 5.5, 5, 5.5), "pt"))
```

**Create education plot**
```{r,echo=FALSE,results=FALSE,warning=FALSE}
# Add a education-wave interaction term to the adjusted model 
education_covar_valid_final<-glmer.nb(mean_mvpa_sd~ wave_id + city_id + wave_id*education + wearRS +
                                                age_cat + employment + income + education +
                                                mean_temp + total_precip + 
                                                (1|interact_id),
                          data=sd_valid[sd_valid$age_cat!="missing" & sd_valid$education!="don't know/prefer not to answer",])

# estimate the effect
education_Inter.HandPick <- effect('wave_id*education', education_covar_valid_final,
                                            xlevels=list(wave_id = c("wave 1","wave 2"),
                                            education = c("university or graduate degree","secondary school",
                                                          "trade school or college")),
                                            se=TRUE, confidence.level=.95, typical=mean)

# put results in data frame 
education_Inter.HandPick <- as.data.frame(education_Inter.HandPick)

# create factor variables for both wave_id and education level          
education_Inter.HandPick$Education <- factor(education_Inter.HandPick$education,
                      levels=c("trade school or college","secondary school","university or graduate degree"),
                      labels=c("College/trade school","High school","University/grad degree"))
education_Inter.HandPick$wave <- factor(education_Inter.HandPick$wave_id,
              levels=c("wave 1","wave 2"),
              labels=c("Wave 1", "Wave 2"))

education_Plot.HandPick<-ggplot(data=education_Inter.HandPick, aes(x=wave, y=fit, group=Education))+
      geom_point(size=1, aes(color=Education),position=position_dodge(0.5))+
      ylim(0,100) +
      geom_errorbar(aes(ymin=lower, ymax=upper,color=Education), width=.3,size=0.6, position=position_dodge(0.5))+
      scale_colour_grey()+
      labs(title="Education level")  +
       guides(color=guide_legend("Education",nrow=4,byrow=TRUE,
                                 keywidth=0.15,
                                 keyheight=0.09,
                                 default.unit="inch"
                                 )) +
                    theme_classic() + theme(plot.title=element_text(size=12,
                                                                     face=2,
                                                                     hjust=0.5),
                                       axis.text.x=element_text(size=10,
                                                              angle = 0,
                                                              face= "bold"),
                                       axis.text.y=element_text(size=10,
                                                              angle = 0),
                                       legend.title=element_blank(),
                                       legend.text = element_text(size=8.5,margin = margin(r = 0.24, unit = 'cm')),
                                       axis.title.x = element_blank(),
                                       axis.title.y = element_blank(),
                                       legend.key.size = unit(0.5,"line"),
                                       legend.position="bottom",
                                       legend.spacing.x = unit(0.06, 'cm'),
                                       legend.margin=margin(-5, 0, 0, 0),
                                       plot.margin = unit(c(12, 5.5, 10, 5.5), "pt"))
```

**Create employment figure**
```{r,echo=FALSE,results=FALSE,warning=FALSE}
# Add an employment status-wave interaction term to the adjusted model 
employment_covar_valid_final<-glmer.nb(mean_mvpa_sd~ wave_id + city_id + wave_id*employment + wearRS +
                                                age_cat + education + income + employment +
                                                mean_temp + total_precip + 
                                                (1|interact_id),
                          data=sd_valid[sd_valid$age_cat!="missing",])

# estimate the effect
employment_Inter.HandPick <- effect('wave_id*employment', employment_covar_valid_final,
                                            xlevels=list(wave_id = c("wave 1","wave 2"),
                                            employment = c("employed","unemployed",
                                                          "retired and not working","other")),
                                            se=TRUE, confidence.level=.95, typical=mean)

# put results in data frame 
employment_Inter.HandPick <- as.data.frame(employment_Inter.HandPick)

# create factor variables for both wave_id and employment status             
employment_Inter.HandPick$Employment <- factor(employment_Inter.HandPick$employment,
                      levels=c("employed","unemployed","retired and not working","other"),
                      labels=c("Employed","Unemployed","Retired","Other"))
employment_Inter.HandPick$wave <- factor(employment_Inter.HandPick$wave_id,
              levels=c("wave 1","wave 2"),
              labels=c("Wave 1", "Wave 2"))

employment_Plot.HandPick<-ggplot(data=employment_Inter.HandPick, aes(x=wave, y=fit, group=Employment))+
      geom_point(size=1, aes(color=Employment),position=position_dodge(0.5))+
      ylim(0,100) +
      geom_errorbar(aes(ymin=lower, ymax=upper,color=Employment), width=.3,size=0.6, position=position_dodge(0.5))+
      scale_colour_grey()+
      labs(title="Employment status")  +
       guides(color=guide_legend("Employment",nrow=4,byrow=TRUE,
                                 keywidth=0.15,
                                 keyheight=0.1,
                                 default.unit="inch"
                                 )) +
                     theme_classic() + theme(plot.title=element_text(size=12,
                                                                     face=2,
                                                                     hjust=0.5),
                                       axis.text.x=element_text(size=10,
                                                              angle = 0,
                                                              face= "bold"),
                                       axis.text.y=element_text(size=10,
                                                              angle = 0),
                                       legend.title=element_blank(),
                                       legend.text = element_text(size=8.5,margin = margin(r = 0.24, unit = 'cm')),
                                       axis.title.x = element_blank(),
                                       axis.title.y = element_blank(),
                                       legend.key.size = unit(0.06,"line"),
                                       legend.position="bottom",
                                       legend.spacing.x = unit(0.1, 'cm'),
                                       legend.margin=margin(-5, 0, 0, 0),
                                       plot.margin = unit(c(12, 5.5, 5, 5.5), "pt"))
```

**Create Figure 3**
```{r,echo=FALSE,warning=FALSE}
titleY=ggpubr::text_grob("Daily minutes of moderate-to-vigorous\nphysical activity",size =12,rot=90,face="bold")

em_figures<-grid.arrange(income_Plot.HandPick,
                         employment_Plot.HandPick,
                         education_Plot.HandPick,
            left=titleY,
            widths=c(1,1,1),
            heights=c(1),
            nrow=1)
ggsave(em_figures,file="figures//figure3.jpeg",width=175,height=85,units="mm")
ggsave(em_figures,file="figures//figure3.pdf",width=175,height=85,units="mm")
ggsave(em_figures,file="submission to IJE//figure3.pdf",width=175,height=85,units="mm")
```

## 4. OBJECTIVE 2: Impact of lockdown stringency on daily MVPA

### A. Main effects

#### Unadjusted model (row 1 table 3)
```{r,echo=FALSE,result=FALSE,warning=FALSE}
  length(sd_pa[sd_pa$wave_id=="wave 2",]$interact_id) # missing 7 observations???
  length(sd_pa[is.na(sd_pa$stringencyIndex) & sd_pa$wave_id=="wave 2",]$interact_id)
```

**Model**
```{r,warning=FALSE}
# model is missing 7 observations from wave 2?
  si_all_unadj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                         data=sd_pa[sd_pa$wave_id=="wave 2",]) 
  summary(si_all_unadj)
```

**Results**
```{r,echo=FALSE,warning=FALSE}
  coefSI<-as.data.frame(summary(si_all_unadj)$coefficients)
  exp(coefSI[2,]$Estimate)
  ci_SI_unadj<-exp(confint(si_all_unadj, c("stringencyIndex"), level = 0.95))
  ci_SI_unadj
  #exp(confint.merMod(si_all_unadj))
```

#### Fully adjusted (row 2 table 3)

**Model**
```{r,warning=FALSE}
  si_all_cov_fully<-glmer.nb(mean_mvpa_sd~ stringencyIndex + wear + city_id + 
                          mean_temp + total_precip + 
                          age_cat + 
                          gender + 
                          children + 
                          employment + 
                          (1|interact_id),
                          data=sd_pa[sd_pa$gender!="other" & sd_pa$age_cat!="missing" &
                                     sd_pa$wave_id=="wave 2",])
  summary(si_all_cov_fully)
```

**Results**
```{r,echo=FALSE,warning=FALSE}
  coefSI<-as.data.frame(summary(si_all_cov_fully)$coefficients)
  exp(coefSI[2,]$Estimate)
  ci_SI_adj<-exp(confint(si_all_cov_fully, c("stringencyIndex"), level = 0.95))
  ci_SI_adj
  #exp(confint.merMod(si_all_cov_fully))
```

### B. Assessing for effect modification

I'll run stratified models to determine if the effect size or direction of the exposure variable (stringencyIndex) varies across levels of each independent variable. These models will be run on valid person-days (600+ minutes of wear time) from wave 2 only.

**Overall, for reference** 
```{r,warning=FALSE}
  si_overall_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id == "wave 2",]) 
  summary(si_overall_valid_adj)
# Comparison: 
  print(paste("overall: ",exp(as.data.frame(summary(si_overall_valid_adj)$coefficients)$Estimate[2])),sep="")
# CIs:
  si_overall_CI<-exp(confint(si_overall_valid_adj, c("stringencyIndex"), level = 0.95))
  print(paste("overall: ",si_overall_CI[1],si_overall_CI[2]),sep="")
```

**Gender**
```{r,echo=FALSE,warning=FALSE}
# models (men, women):
  si_gender1_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex  + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id == "wave 2" & sd_valid$gender=="man",]) 
  si_gender2_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex  + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id == "wave 2" & sd_valid$gender=="woman",])

# effect estimates:
  si_gender1_coef<-round(exp(as.data.frame(summary(si_gender1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_gender2_coef<-round(exp(as.data.frame(summary(si_gender2_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  si_gender1_CI<-round(exp(confint(si_gender1_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_gender2_CI<-round(exp(confint(si_gender2_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
```

Distribution of person-days by gender
```{r,echo=FALSE}
  addmargins(table(as.factor(sd_valid[sd_valid$wave_id=="wave 2",]$gender)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
  print(paste("men: ",si_gender1_coef[1]," (",si_gender1_CI[1],", ",si_gender1_CI[2],")",sep=""))
  print(paste("women: ",si_gender2_coef[1]," (",si_gender2_CI[1],", ",si_gender2_CI[2],")",sep=""))
```

**Age** 
```{r,echo=FALSE,warning=FALSE}
# models (18-29, 30-49, 50-69, 70+) 
  si_age1_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$age_cat=="18-29",]) 
  si_age2_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$age_cat=="30-49",]) 
  si_age3_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$age_cat=="50-69",]) 
  si_age4_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$age_cat=="70+",])

# effect estimates:
  si_age1_coef<-round(exp(as.data.frame(summary(si_age1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_age2_coef<-round(exp(as.data.frame(summary(si_age2_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_age3_coef<-round(exp(as.data.frame(summary(si_age3_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_age4_coef<-round(exp(as.data.frame(summary(si_age4_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  si_age1_CI<-round(exp(confint(si_age1_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_age2_CI<-round(exp(confint(si_age2_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_age3_CI<-round(exp(confint(si_age3_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_age4_CI<-round(exp(confint(si_age4_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
```

Distribution of person-days by age
```{r,echo=FALSE}
addmargins(table(as.factor(sd_valid[sd_valid$wave_id=="wave 2",]$age_cat)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
  print(paste("18-29: ",si_age1_coef[1]," (",si_age1_CI[1],", ",si_age1_CI[2],")",sep=""))
  print(paste("30-49: ",si_age2_coef[1]," (",si_age2_CI[1],", ",si_age2_CI[2],")",sep=""))
  print(paste("50-69: ",si_age3_coef[1]," (",si_age3_CI[1],", ",si_age3_CI[2],")",sep=""))
  print(paste("70+: ",   si_age4_coef[1]," (",si_age4_CI[1],", ",si_age4_CI[2],")",sep=""))
```

**Income**
```{r,echo=FALSE,warning=FALSE}
# models (<50,000; 50-99,999; 100,000+; don't know/prefer not to answer) 
  si_income1_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$income=="<$50,000",]) 
  si_income2_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$income=="$50,000 to $99,999",]) 
  si_income3_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$income=="$100,000 or more",]) 
  si_income4_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$income=="don't know/prefer not to answer",])

# effect estimates:
  si_income1_coef<-round(exp(as.data.frame(summary(si_income1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_income2_coef<-round(exp(as.data.frame(summary(si_income2_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_income3_coef<-round(exp(as.data.frame(summary(si_income3_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_income4_coef<-round(exp(as.data.frame(summary(si_income4_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  si_income1_CI<-round(exp(confint(si_income1_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_income2_CI<-round(exp(confint(si_income2_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_income3_CI<-round(exp(confint(si_income3_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_income4_CI<-round(exp(confint(si_income4_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
```

Distribution of person-days by income
```{r,echo=FALSE}
  addmargins(table(as.factor(sd_valid[sd_valid$wave_id=="wave 2",]$income)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
  print(paste("<$50,000: ",si_income1_coef[1]," (",si_income1_CI[1],", ",si_income1_CI[2],")",sep=""))
  print(paste("$50,000 to $99,999: ",si_income2_coef[1]," (",si_income2_CI[1],", ",si_income2_CI[2],")",sep=""))
  print(paste("$100,000 or more: ",si_income3_coef[1]," (",si_income3_CI[1],", ",si_income3_CI[2],")",sep=""))
  print(paste("don't know/perfer not to answer: ",   si_income4_coef[1]," (",si_income4_CI[1],", ",si_income4_CI[2],")",sep=""))
```

**Education**
```{r,echo=FALSE,warning=FALSE}
# models (university or grad degree, secondary school, trade school or college) 
  si_education1_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$education=="university or graduate degree",]) 
  si_education2_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$education=="secondary school",]) 
  si_education3_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$education=="trade school or college",])

# effect estimates:
  si_education1_coef<-round(exp(as.data.frame(summary(si_education1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_education2_coef<-round(exp(as.data.frame(summary(si_education2_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_education3_coef<-round(exp(as.data.frame(summary(si_education3_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  si_education1_CI<-round(exp(confint(si_education1_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_education2_CI<-round(exp(confint(si_education2_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_education3_CI<-round(exp(confint(si_education3_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
```

Distribution of person-days by education
```{r,echo=FALSE}
  addmargins(table(as.factor(sd_valid[sd_valid$wave_id=="wave 2",]$education)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
  print(paste("university or grad degree: ",si_education1_coef[1]," (",si_education1_CI[1],", ",si_education1_CI[2],")",sep=""))
  print(paste("secondary school: ",si_education2_coef[1]," (",si_education2_CI[1],", ",si_education2_CI[2],")",sep=""))
  print(paste("trade school or college: ",si_education3_coef[1]," (",si_education3_CI[1],", ",si_education3_CI[2],")",sep=""))
```

**Employment status** 
```{r,echo=FALSE,warning=FALSE}
# models (employed, retired and not working, unemployed, other) 
  si_employment1_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$employment=="employed",]) 
  si_employment2_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$employment=="retired and not working",]) 
  si_employment3_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$employment=="unemployed",]) 
  si_employment4_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$employment=="other",])

# effect estimates:
  si_employment1_coef<-round(exp(as.data.frame(summary(si_employment1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_employment2_coef<-round(exp(as.data.frame(summary(si_employment2_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_employment3_coef<-round(exp(as.data.frame(summary(si_employment3_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_employment4_coef<-round(exp(as.data.frame(summary(si_employment4_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  si_employment1_CI<-round(exp(confint(si_employment1_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_employment2_CI<-round(exp(confint(si_employment2_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_employment3_CI<-round(exp(confint(si_employment3_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_employment4_CI<-round(exp(confint(si_employment4_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
```

Distribution of person-days by employement status
```{r,echo=FALSE}
 addmargins(table(as.factor(sd_valid[sd_valid$wave_id=="wave 2",]$employment)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
print(paste("employed: ",si_employment1_coef[1]," (",si_employment1_CI[1],", ",si_employment1_CI[2],")",sep=""))
  print(paste("retired and not working: ",si_employment2_coef[1]," (",si_employment2_CI[1],", ",si_employment2_CI[2],")",sep=""))
  print(paste("unemployed: ",si_employment3_coef[1]," (",si_employment3_CI[1],", ",si_employment3_CI[2],")",sep=""))
  print(paste("other: ",   si_employment4_coef[1]," (",si_employment4_CI[1],", ",si_employment4_CI[2],")",sep=""))
```

**Has children**
```{r,echo=FALSE,warning=FALSE}
# models (has children living at home, does not have children living at home) 
  si_children1_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$children=="yes",]) 
  si_children2_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$children=="no",])

# effect estimates:
  si_children1_coef<-round(exp(as.data.frame(summary(si_children1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_children2_coef<-round(exp(as.data.frame(summary(si_children2_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  si_children1_CI<-round(exp(confint(si_children1_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_children2_CI<-round(exp(confint(si_children2_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
```

Distribution of person-days by employment status
```{r,echo=FALSE}
  addmargins(table(as.factor(sd_valid[sd_valid$wave_id=="wave 2",]$children)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
  print(paste("has children at home: ",si_children1_coef[1]," (",si_children1_CI[1],", ",si_children1_CI[2],")",sep=""))
  print(paste("does not have children at home: ",si_children2_coef[1]," (",si_children2_CI[1],", ",si_children2_CI[2],")",sep=""))
```

**City**
```{r,echo=FALSE,warning=FALSE}
# models (montreal, saskatoon, vancouver) 
  si_city1_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$city=="montreal",]) 
  si_city2_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$city=="saskatoon",]) 
  si_city3_valid_adj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
                              data=sd_valid[sd_valid$wave_id=="wave 2" & sd_valid$city=="vancouver",])

# effect estimates:
  si_city1_coef<-round(exp(as.data.frame(summary(si_city1_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_city2_coef<-round(exp(as.data.frame(summary(si_city2_valid_adj)$coefficients)$Estimate[2]),digits=2)
  si_city3_coef<-round(exp(as.data.frame(summary(si_city3_valid_adj)$coefficients)$Estimate[2]),digits=2)

# CIs:
  set.seed(123)
  si_city1_CI<-round(exp(confint(si_city1_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_city2_CI<-round(exp(confint(si_city2_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
  si_city3_CI<-round(exp(confint(si_city3_valid_adj, c("stringencyIndex"), level = 0.95,method="boot")),digits=2)
```

Distribution of person-days by city
```{r,echo=FALSE}
  addmargins(table(as.factor(sd_valid[sd_valid$wave_id=="wave 2",]$city)))
```

Effect estimate and CI for each strata
```{r,echo=FALSE}
  print(paste("montreal: ",si_city1_coef[1]," (",si_city1_CI[1],", ",si_city1_CI[2],")",sep=""))
  print(paste("saskatoon: ",si_city2_coef[1]," (",si_city2_CI[1],", ",si_city2_CI[2],")",sep=""))
  print(paste("vancouver: ",si_city3_coef[1]," (",si_city3_CI[1],", ",si_city3_CI[2],")",sep=""))
```

## 5. Supplementary Figure 1 (flowchart)

*Objective 1*

Full sample
```{r}
length(sd_pa$date)
```

Unadjusted model (all person-days with 600+ minutes of wear time)
```{r}
length(sd_valid$date)
table(sd_valid$wave_id)
```
Adjusted model (complete case; no missing data for confounders)
```{r}
# models adjusted for wave_id + city_id + wearRS + age_cat + employment + income + mean_temp + total_precip -- no missing data for wave_id, city_id, employment status, income... look for number of missing for each of the remaining variables

print("age")
summary(sd_valid$age_cat)

print("mean_temp")
summary(sd_valid$mean_temp)

print("total_precip")
summary(sd_valid$total_precip)

# any with multiple missing?
length(sd_valid[is.na(sd_valid$total_precip) & is.na(sd_valid$mean_temp),]$date) # 88 missing temp are also missing precip and vice versa
length(sd_valid[is.na(sd_valid$age_cat) & is.na(sd_valid$mean_temp),]$date) # no overlap between missing temp/precip and age

length(sd_valid[!is.na(sd_valid$age_cat) & !is.na(sd_valid$mean_temp),]$date)

table(sd_valid[!is.na(sd_valid$age_cat) & !is.na(sd_valid$mean_temp),]$wave_id)
```

*Sensitivity analysis 1*

Unadjusted
```{r}
length(sd_valid_both$date)
```

Adjusted
```{r}
print("age")
summary(sd_valid_both$age_cat)

print("mean_temp")
summary(sd_valid_both$mean_temp)

print("total_precip")
summary(sd_valid_both$total_precip)
```
*Sensitivity analysis 2*

Unadjusted
```{r}
length(sens_valid$date)
```

Adjusted
```{r}
print("age")
summary(sens_valid$age_cat) # don't need to check for missingness in weather data since we aren't controlling for weather in this model
```

*Objective 2*

Unadjusted model (all wave 2 person-days with 600+ minutes of wear time)
```{r}
length(sd_valid[sd_valid$wave_id=="wave 2",]$date)
```

Adjusted model (complete case; no missing data for confounders)
```{r}
print("age")
summary(sd_valid[sd_valid$wave_id=="wave 2",]$mean_temp)
```

### 2023-10-11 paper1-INTERACT-data-prep.Rmd
Trying to create a public repo for the INTERACT analysis

---
title: "paper1-wrangling"
author: "Shelby Sturrock"
date: "26/01/2022"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE,results=FALSE,echo=FALSE}
  knitr::opts_chunk$set(echo = TRUE)
  library(dplyr)
  library(lubridate)
  library(ggplot2)
  library(tableone)
  library(table1)
  library(tidyr)
  library(lme4)
  library(gridExtra)
  knitr::opts_knit$set(root.dir = '/Users/shelbysturrock/Library/Mobile Documents/com~apple~CloudDocs/Documents/Documents - Shelby’s MacBook Pro/School/UofT/Dissertation/PAPER 1 - INTERACT')
```

# Paper 1: INTERACT

## 1. Read in and process data

### Sensdoc data

Objective physical activity as measured by research-grade accelerometers provided to participants.

Read in SD data: 
```{r,results="hide"}
  sensdoc<-list.files("Data/sensdoc data", pattern="*.csv", full.names=TRUE)
  sensdoc_csv<-lapply(sensdoc, read.csv)
  names(sensdoc_csv)<-paste(gsub("Data/sensdoc data/|.csv|\\_top.*","",sensdoc),"_sd",sep="")
  names(sensdoc_csv)<-ifelse(grepl("saskatoon",names(sensdoc_csv)),gsub("saskatoon","skt",names(sensdoc_csv)),
                             ifelse(grepl("montreal",names(sensdoc_csv)),gsub("montreal","mtl",names(sensdoc_csv)),
                                                                         gsub("vancouver","van",names(sensdoc_csv))))
  names(sensdoc_csv)<-gsub("_0","",names(sensdoc_csv))
  list2env(sensdoc_csv,envir=.GlobalEnv)
  rm(sensdoc,sensdoc_csv)
```

Reformat date variables
```{r}
# wave 1
  mtl1_sd$date<-as.Date(mtl1_sd$utcdate,"%Y-%m-%d")
  van1_sd$date<-as.Date(van1_sd$utcdate,"%Y-%m-%d")
  skt1_sd$date<-as.Date(skt1_sd$utcdate,"%Y-%m-%d")
# wave 2
  mtl2_sd$date<-as.Date(mtl2_sd$utcdate,"%Y-%m-%d")
  van2_sd$date<-as.Date(van2_sd$utcdate,"%Y-%m-%d")
  skt2_sd$date<-as.Date(skt2_sd$utcdate,"%Y-%m-%d")
```

### Eligibility/survey data

Data from the eligibility survey.

```{r}
  eligib<-list.files("/Users/shelbysturrock/Library/Mobile Documents/com~apple~CloudDocs/Documents/Documents - Shelby’s MacBook Pro/School/UofT/Dissertation/PAPER 1 - INTERACT/Data/surveys/eligibility", pattern="*.csv", full.names=TRUE)
  eligib_csv<-lapply(eligib, read.csv)
  names(eligib_csv)<-paste(gsub("/Users/shelbysturrock/Library/Mobile Documents/com~apple~CloudDocs/Documents/Documents - Shelby’s MacBook Pro/School/UofT/Dissertation/PAPER 1 - INTERACT/Data/surveys/eligibility/eligibility_|.csv|\\_main.*","",eligib),"_eligib",sep="")
  names(eligib_csv)<-ifelse(grepl("1",names(eligib_csv)),gsub("_","1_",gsub("1","",names(eligib_csv))),
                     ifelse(grepl("new_",names(eligib_csv)),paste(gsub("_","2_",gsub("2|new|","",names(eligib_csv))),"_new",sep=""),
                            gsub("_","2_",gsub("2","",names(eligib_csv)))))
# names(eligib_csv)<-ifelse(grepl("1",names(eligib_csv)),paste(gsub("1","",names(eligib_csv)),"1",sep=""),
#                                                        paste(gsub("2","",names(eligib_csv)),"2",sep=""))
  list2env(eligib_csv,envir=.GlobalEnv)
  rm(eligib,eligib_csv)
```

### Health survey

Read in health survey data
```{r}
  health<-list.files("/Users/shelbysturrock/Library/Mobile Documents/com~apple~CloudDocs/Documents/Documents - Shelby’s MacBook Pro/School/UofT/Dissertation/PAPER 1 - INTERACT/Data/surveys/health", pattern="*.csv", full.names=TRUE)
  health_csv<-lapply(health, read.csv)
  names(health_csv)<-paste(gsub("/Users/shelbysturrock/Library/Mobile Documents/com~apple~CloudDocs/Documents/Documents - Shelby’s MacBook Pro/School/UofT/Dissertation/PAPER 1 - INTERACT/Data/surveys/health/health_|.csv|\\_main.*","",health),"_health",sep="")
  names(health_csv)<-ifelse(grepl("1",names(health_csv)),gsub("_","1_",gsub("1","",names(health_csv))),
                     ifelse(grepl("new_",names(health_csv)),paste(gsub("_","2_",gsub("2|new|","",names(health_csv))),"_new",sep=""),
                            gsub("_","2_",gsub("2","",names(health_csv)))))
  list2env(health_csv,envir=.GlobalEnv)
  rm(health,health_csv)
```

Health columns of interest and combine health survey data
```{r}
# wave 1 
  skt1hlt<-skt1_health %>% dplyr::select(interact_id,wave_id,gender,children,income,marital_status,dwelling_type,employment,education)
  mtl1hlt<-mtl1_health %>% dplyr::select(colnames(skt1hlt))
  van1hlt<-van1_health %>% dplyr::select(colnames(skt1hlt))

# wave  2
  skt2hlt<-skt2_health %>% dplyr::select(colnames(skt1hlt))
  mtl2hlt<-mtl2_health %>% dplyr::select(colnames(skt1hlt))
  van2hlt<-van2_health %>% dplyr::select(colnames(skt1hlt))

# wave 2 new
  skt2hlt_new<-skt2_health_new %>% dplyr::select(colnames(skt1hlt))
  mtl2hlt_new<-mtl2_health_new %>% dplyr::select(colnames(skt1hlt))
  van2hlt_new<-van2_health_new %>% dplyr::select(colnames(skt1hlt))

# combine health surveys into one data frame
  health<-rbind(skt1hlt,mtl1hlt,van1hlt,
                skt2hlt,mtl2hlt,van2hlt,
                skt2hlt_new,mtl2hlt_new,van2hlt_new)
  rm(list=ls()[grepl("_health|hlt",ls())])
```

## 2. Process/clean demographic, health data; merge with SD data

### Age: calculate using data from from the eligibility survey

#### Saskatoon: Review data

Assess for missing ages - Saskatoon, wave 1
```{r}
length(skt1_eligib[skt1_eligib$birth_date=="",]$interact_id)
```
18 participants have missing date of birth in the wave 1 eligibility survey.

Assess for missing ages - Saskatoon, wave 2
```{r}
length(skt2_eligib[skt2_eligib$birth_date=="",]$interact_id)
```
0 participants are missing date of brith in the wave 2 eligibility survey.

Is everyone in the wave 1 eligbility survey also in wave 2 eligibility survey? If so, we can steal DOB from the wave 2 eligibility survey to assess age in wave 1.

```{r}
length(skt2_eligib[!(skt2_eligib$interact_id %in% skt1_eligib$interact_id),]$interact_id)
```

Confirmed: everyone in skt_eligb2 is in skt1_eligib -- if any of these participants have a missing DOB in wave 1, it can be determined based on wave 2 survey.

Since there are no missing DOB in the wave 2 eligibility survey, link the wave 1 and wave 2 eligibility survey to update DOB for those people whose age is NA (missing DOB or aged < 18)

For wave 1:

  Calculate age using date_of_birth in eligib1, where available.

  Where NOT available, pull birth_date from eligib2.

Calculate age wave 2:

  Update age for those people in wave 1 and wave 2 (mtl2_eligib).

  Calculate age for those people new in wave 2 (mtl2_eligib_new).

#### Saskatoon: Calculate age in wave 1

Extract interact_id and birth_date from wave 2 eligibility survey
```{r}
  skt_age2<-skt2_eligib %>% dplyr::select(interact_id,birth_date)
```

Re-code birth_date to missing if cell is empty or contains invalid DOB
```{r}
  skt1_eligib[skt1_eligib$birth_date =="" | grepl("0080",skt1_eligib$birth_date),]$birth_date<-NA
```

If birth_date is NOT missing, calculate age as "date_of_survey" - "birth_date", otherwise set age as missing. Also set as missing if age is < 18
```{r}
  skt1_eligib$age<-ifelse(!is.na(skt1_eligib$birth_date),
                          round((difftime(as.Date(skt1_eligib$date_of_survey,origin="1970-01-01"),                                                       as.Date(skt1_eligib$birth_date,origin="1970-01-01"),unit="weeks")/52.25),0),NA)
  summary(skt1_eligib$age)
  skt1_eligib[!is.na(skt1_eligib$age) & skt1_eligib$age < 18,]$age<-NA
  skt1_eligib<-skt1_eligib %>% dplyr::select(interact_id,age,date_of_survey,birth_date)
```

Merge with birth_date from eligib2, by interact_id, to populate birth_date where it is currently missing
```{r}
  skt1_eligib<-merge(skt1_eligib,skt_age2,by="interact_id",all.x=TRUE)
```

Where age is missing in eligib1, and it is NOT missing in eligib2, calculate age as "date_of_survey" - "birth_date"
```{r}
  skt1_eligib$age<-ifelse(is.na(skt1_eligib$age) & !is.na(skt1_eligib$birth_date.y),
                          round((difftime(as.Date(skt1_eligib$date_of_survey,origin="1970-01-01"),
                          as.Date(skt1_eligib$birth_date.y,origin="1970-01-01"),unit="weeks")/52.25),0),skt1_eligib$age) 
  length(skt1_eligib[is.na(skt1_eligib$age),]$interact_id)
    # decreased NA age from 86 records to 67 records (reduction of 19 -- the number of people missing DOB in wave 1 but DOB populated in wave 2)
```

Merge age into the wave 1 sd data
```{r}
    skt1_sd<-merge(skt1_sd[,colnames(skt1_sd)!="age"],skt1_eligib,by="interact_id",all.x=TRUE)
```

#### Saskatoon: Calculate age in wave 2

Age for wave 2 - calculate age at time of survey completion based on eligibility survey.

Start by simplifying the wave 2 eligibility data (just interact_id, birth_date, date_of_survey) 
```{r}
  skt2_eligib<-skt2_eligib %>% dplyr::select(interact_id,birth_date,date_of_survey)
```

Pull the same columns from "new" eligibility survey (presumably new participants?)
```{r}
  skt2_eligib_new<-skt2_eligib_new %>% dplyr::select(colnames(skt2_eligib))
```

Create a list of participants with data in the wave 2 eligibility surveys (for existing and new participants) and then
```{r}
  skt2_eligib<-distinct(rbind(skt2_eligib,skt2_eligib_new))
```

Calculate age at time of survey completion 
```{r}
  skt2_eligib$age<-ifelse(!is.na(skt2_eligib$birth_date),
                          round((difftime(as.Date(skt2_eligib$date_of_survey,origin="1970-01-01"),                                                                       as.Date(skt2_eligib$birth_date,origin="1970-01-01"),unit="weeks")/52.25),0),NA)
```

Replace implausibly small ages with NA
```{r}
  skt2_eligib[!is.na(skt2_eligib$age) & skt2_eligib$age < 18,]$age<-NA
  skt2_eligib<-skt2_eligib %>% dplyr::select(interact_id,age)
  skt1_eligib<-skt1_eligib %>% dplyr::select(interact_id,age)
```

Merge age into the wave 2 sd data
```{r}
  skt2_sd<-merge(skt2_sd,skt2_eligib,by="interact_id",all.x=TRUE)
```

Note: 36 people (everyoen with a missing DOB/age in wave 2) are NOT in either the wave 1 or wave 2 eligibility survey
```{r}
  length(unique(skt2_sd[is.na(skt2_sd$age),]$interact_id))
```

#### Montreal: Review data

```{r}
  length(mtl1_eligib[mtl1_eligib$birth_date=="",]$interact_id) # no missing DOB
```

```{r}
  summary(as.factor(mtl2_eligib$birth_date)) # no DOB column
```

```{r}
  length(mtl2_eligib$interact_id %in% mtl1_eligib$interact_id) # confirmed: everyone in mtl_eligb2 is in mtl1_eligib   
  length(mtl1_eligib[mtl2_eligib_new$birth_date=="",]$interact_id) # no missing DOB
```

  Calculate age for everyone in wave 1 -- "date_of_survey" - "birth_date"

  Calculate age in wave 2:

   Update age for those people in wave 1 and wave 2 (mtl2_eligib)

   Calculate age for those people new in wave 2 (mtl2_eligib_new)

#### Montreal: Calculate age in wave 1

There are no participants with a missing DOB. As a result, age can be calculated as "date_of_survey" minus "birth_date" for all participants
```{r}
  mtl1_eligib$age<-ifelse(!is.na(mtl1_eligib$birth_date),
                          round((difftime(as.Date(mtl1_eligib$date_of_survey,origin="1970-01-01"),
                                          as.Date(mtl1_eligib$birth_date,origin="1970-01-01"),unit="weeks")/52.25),0),NA)
# select ID, age
  #mtl_age1_merge<-mtl1_eligib %>% dplyr::select(interact_id,age)
```

Merge age into the wave 1 sd data
```{r}
# merge age with wave 1 sd data
  mtl1_sd<-merge(mtl1_sd[,colnames(mtl1_sd)!="age"],
                 mtl1_eligib[,colnames(mtl1_eligib) %in% c("interact_id","age")],
                 by="interact_id",all.x=TRUE)
```

#### Montreal: Calculate age in wave 2

Pull birth_date from wave 1 eligibility survey; merge with wave 2
```{r}
  mtl_age1<-mtl1_eligib %>% dplyr::select(interact_id,birth_date)
  mtl2_eligib<-merge(mtl2_eligib,mtl_age1,by="interact_id")
```

Start by simplifying the wave 2 eligibility data (just interact_id, birth_date, date_of_survey) 
```{r}
 mtl2_eligib<-mtl2_eligib %>% dplyr::select(interact_id,birth_date,date_of_survey)
```

Pull the same columns from "new" eligibility survey (presumably new participants?)
```{r}
 mtl2_eligib_new<-mtl2_eligib_new %>% dplyr::select(colnames(mtl2_eligib))
```

Create a list of participants with data in the wave 2 eligibility surveys (for existing and new participants) and then
```{r}
 mtl2_eligib<-distinct(rbind(mtl2_eligib,mtl2_eligib_new))
```

Calculate age at time of survey completion 
```{r}
  mtl2_eligib$age<-ifelse(!is.na(mtl2_eligib$birth_date),
                          round((difftime(as.Date(mtl2_eligib$date_of_survey,origin="1970-01-01"),
                                          as.Date(mtl2_eligib$birth_date,origin="1970-01-01"),unit="weeks")/52.25),0),NA)
```

Merge age with wave 2 sd data
```{r}
  mtl2_sd<-merge(mtl2_sd[,colnames(mtl2_sd)!="age"],mtl2_eligib,by="interact_id",all.x=TRUE)
```

#### Vancouver: Review data

```{r}
  length(van1_eligib[van1_eligib$birth_date=="",]$interact_id) # no missing DOB
```

```{r}
  summary(as.factor(van2_eligib$birth_date)) # no DOB column
```

```{r}
  length(van2_eligib$interact_id %in% van1_eligib$interact_id) # confirmed: everyone in van_eligb2 is in van1_eligib 
  length(mtl1_eligib[van2_eligib_new$birth_date=="",]$interact_id) # no missing DOB
```

 Calculate age for everyone in wave 1 -- "date_of_survey" - "birth_date"

 Calculate age in wave 2:

   Update age for those people in wave 1 and wave 2 (van2_eligib)

   Calculate age for those people new in wave 2 (van2_eligib_new)

#### Vancouver: Calculate age in wave 1

There are no participants with a missing DOB. As a result, age can be calculated as "date_of_survey" minus "birth_date" for all participants
```{r}
  van1_eligib$age<-ifelse(!is.na(van1_eligib$birth_date),
                          round((difftime(as.Date(van1_eligib$date_of_survey,origin="1970-01-01"),
                                          as.Date(van1_eligib$birth_date,origin="1970-01-01"),unit="weeks")/52.25),0),NA)

  #van_age1_merge<-van1_eligib %>% dplyr::select(interact_id,age)
```

Merge age into the wave 1 sd data
```{r}
  van1_sd<-merge(van1_sd[,colnames(van1_sd)!="age"],
                 van1_eligib[,colnames(van1_eligib) %in% c("interact_id","age")],
                 by="interact_id",all.x=TRUE)
```

#### Vancouver: Calculate age in wave 2

Pull birth_date from wave 1 eligibility survey; merge with wave 2
```{r}
  van_age1<-van1_eligib %>% dplyr::select(interact_id,birth_date)
  van2_eligib<-merge(van2_eligib,van_age1,by="interact_id")
```

Start by simplifying the wave 2 eligibility data (just interact_id, birth_date, date_of_survey) 
```{r}
  van2_eligib<-van2_eligib %>% dplyr::select(interact_id,birth_date,date_of_survey)
```

Pull the same columns from "new" eligibility survey (presumably new participants?)
```{r}
  van2_eligib_new<-van2_eligib_new %>% dplyr::select(colnames(van2_eligib))
```

Create a list of participants with data in the wave 2 eligibility surveys (for existing and new participants) and then
```{r}
  van2_eligib<-distinct(rbind(van2_eligib,van2_eligib_new))
```

Calculate age at time of survey completion 
```{r}
  van2_eligib$age<-ifelse(!is.na(van2_eligib$birth_date),
                          round((difftime(as.Date(van2_eligib$date_of_survey,origin="1970-01-01"),
                                          as.Date(van2_eligib$birth_date,origin="1970-01-01"),unit="weeks")/52.25),0),NA)

```

Merge age with wave 2 sd data
```{r}
  van2_sd<-merge(van2_sd[,colnames(van2_sd)!="age"],van2_eligib,by="interact_id",all.x=TRUE)
```

Remove eligibility datasets
```{r}
  rm(van1_eligib,van2_eligib,van2_eligib_new,skt1_eligib,skt2_eligib,skt2_eligib_new,mtl1_eligib,mtl2_eligib,mtl2_eligib_new,
     mtl_age1,skt_age2,van_age1)
```

## 3. Prep SD data

### Combine all SD data into single dataframe

Combine SD data into a single dataframe: 
```{r}
# wave 1
  mtl1_sd<-mtl1_sd %>% dplyr::select(interact_id,utcdate,lat,lon,easting,northing,zone,in_city,
                                     x_count,y_count,z_count,summary_count,wearing,activity_levels,date,age,
                                     wave_id,city_id)
  van1_sd<-van1_sd %>% dplyr::select(colnames(mtl1_sd))
  skt1_sd<-skt1_sd %>% dplyr::select(colnames(mtl1_sd))

# wave 2
  mtl2_sd<-mtl2_sd %>% dplyr::select(colnames(mtl1_sd))
  van2_sd<-van2_sd %>% dplyr::select(colnames(mtl2_sd))
  skt2_sd<-skt2_sd %>% dplyr::select(colnames(mtl2_sd))

# create a combined datasets
  sd<-rbind(skt1_sd,mtl1_sd,van1_sd,
            skt2_sd,mtl2_sd,van2_sd)
```

Add age_cat variable
```{r}
  sd$age_cat<-ifelse(!is.na(sd$age) & sd$age < 30,"18-29",
              ifelse(!is.na(sd$age) & sd$age >= 30 & sd$age < 50,"30-49",
              ifelse(!is.na(sd$age) & sd$age >= 50 & sd$age < 70,"50-69",
              ifelse(!is.na(sd$age) & sd$age >= 70,"70+","missing"))))
```

Test: wearing by wave_id and in_city by wave_id
```{r}
print("wearing:")
table(sd$wearing,sd$wave_id)
print(" ")
print("in_city:")
table(sd$in_city,sd$wave_id)
```

### Aggregate PA data at the level of the person-day

*NEW*: Create new activity_levels variable using logic from the wave 1 baseline paper (applied to x-count, as in the baseline paper)
```{r}
  sd<-sd %>% mutate(activity_levels_shelby =  case_when(x_count < 100 ~ "sedentary",
                                                  		  x_count >= 100 & x_count <= 1951 ~ "light",
                                                  		  x_count >= 1951 & x_count <= 5724 ~ "moderate",
                                                  	    x_count >= 5725 ~ "vigorous"))
```

Remove individual city-wave dataframes
```{r}
  rm(skt1_sd,mtl1_sd,van1_sd,
     skt2_sd,mtl2_sd,van2_sd)
```

Create binary indicators for MVPA, sedentary and LPA using (a) activity_levels ("Old", from the compute canada data) and (b) activity_levels_shelby (calculated above using threshold pulled from Dan's baseline paper code)
```{r}
  sd$mvpaOld<-ifelse(sd$activity_levels %in% c("moderate","vigorous"),1,0)
  sd$lightOld<-ifelse(sd$activity_levels=="light",1,0)
  sd$sbOld<-ifelse(sd$activity_levels=="sedentary",1,0)

  sd$mvpa<-ifelse(sd$activity_levels_shelby %in% c("moderate","vigorous"),1,0)
  sd$light<-ifelse(sd$activity_levels_shelby == "light",1,0)
  sd$sb<-ifelse(sd$activity_levels_shelby == "sedentary",1,0)
```

*NEW*: Create 3 versions of the SD data to work from: 
1. All minutes
2. Minutes where wearing == 1
3. Minutes where wearing == 1 & in_city == 1
```{r}
# 1. all minutes
  sd_all<-sd
# 2. minutes where wearing == 1
  wearing<-sd[sd$wearing==1,]
# 3. minutes where wearing == 1 & in_city == 1
  wearingCity<-sd[sd$in_city==1 & sd$wearing==1,]
```

Add additional summary measures
```{r}
# 1. all person days
  sd_all<-sd_all %>% group_by(interact_id,date,city_id,wave_id) %>% 
             mutate(minutes=length(date),
                    wear=sum(wearing),
                # using original activity_levels:
                    total_mvpa_minutesOld=sum(mvpaOld),
                    total_light_pa_minutesOld=sum(lightOld),
                    total_sed_minutesOld=sum(sbOld),
                # using updated activity_levels_shelby: 
                    total_mvpa_minutes=sum(mvpa),
                    total_light_pa_minutes=sum(light),
                    total_sed_minutes=sum(sb))

# 2. person days where wearing == 1
  sd_wearing<-wearing %>% group_by(interact_id,date,city_id,wave_id) %>% 
             mutate(minutes=length(date),
                    wear=sum(wearing),
                # using original activity_levels:
                    total_mvpa_minutesOld=sum(mvpaOld),
                    total_light_pa_minutesOld=sum(lightOld),
                    total_sed_minutesOld=sum(sbOld),
                # using updated activity_levels_shelby: 
                    total_mvpa_minutes=sum(mvpa),
                    total_light_pa_minutes=sum(light),
                    total_sed_minutes=sum(sb))

# 3. person-days with wearing == 1 and in_city == 1
  sd_wearingCity<-wearingCity %>% group_by(interact_id,date,city_id,wave_id) %>% 
             mutate(minutes=length(date),
                    wear=sum(wearing),
                # using original activity_levels:
                    total_mvpa_minutesOld=sum(mvpaOld),
                    total_light_pa_minutesOld=sum(lightOld),
                    total_sed_minutesOld=sum(sbOld),
                # using updated activity_levels_shelby: 
                    total_mvpa_minutes=sum(mvpa),
                    total_light_pa_minutes=sum(light),
                    total_sed_minutes=sum(sb))
```

Group at the individual level 
```{r,warning=FALSE,message=FALSE}
# 1. all person-days
  sd_all <- sd_all %>% group_by(interact_id, date, city_id, wave_id) %>%
    summarize(
      # wear time
        wearing = mean(wearing, na.rm = TRUE), # proportion of time while wearing
        minutes = mean(minutes, na.rm = TRUE),
        wear = mean(wear, na.rm = TRUE), # minutes of wear time
      # activity using activity_levels_shelby:
        mean_mvpa_sdOld = mean(total_mvpa_minutesOld, na.rm = TRUE),
        mean_sed_sdOld = mean(total_sed_minutesOld, na.rm = TRUE), 
        mean_light_sdOld = mean(total_light_pa_minutesOld, na.rm = TRUE), 
      # activity using original activity_levels: 
        mean_mvpa_sd = mean(total_mvpa_minutes, na.rm = TRUE),
        mean_sed_sd = mean(total_sed_minutes, na.rm = TRUE), 
        mean_light_sd = mean(total_light_pa_minutes, na.rm = TRUE), 
      # demographics: 
        age = mean(age),
        age_cat = identity(age_cat))
  sd_all <- distinct(sd_all)

# 2. person-days where wearing == 1
  sd_wearing <- sd_wearing %>% group_by(interact_id, date, city_id, wave_id) %>%
    summarize(
      # wear time
        wearing = mean(wearing, na.rm = TRUE), # proportion of time while wearing
        minutes = mean(minutes, na.rm = TRUE),
        wear = mean(wear, na.rm = TRUE), # minutes of wear time
      # activity using original activity_levels: 
        mean_mvpa_sd = mean(total_mvpa_minutes, na.rm = TRUE),
        mean_sed_sd = mean(total_sed_minutes, na.rm = TRUE), 
        mean_light_sd = mean(total_light_pa_minutes, na.rm = TRUE), 
      # demographics: 
        age = mean(age),
        age_cat = identity(age_cat))
  sd_wearing <- distinct(sd_wearing)

# 3. person-days where in_city == 1 and wearing == 1
  sd_wearingCity <- sd_wearingCity %>% group_by(interact_id, date, city_id, wave_id) %>%
    summarize(
      # wear time
        wearing = mean(wearing, na.rm = TRUE), # proportion of time while wearing
        minutes = mean(minutes, na.rm = TRUE),
        wear = mean(wear, na.rm = TRUE), # minutes of wear time
      # activity using original activity_levels: 
        mean_mvpa_sd = mean(total_mvpa_minutes, na.rm = TRUE),
        mean_sed_sd = mean(total_sed_minutes, na.rm = TRUE), 
        mean_light_sd = mean(total_light_pa_minutes, na.rm = TRUE), 
      # demographics: 
        age = mean(age),
        age_cat = identity(age_cat))
  sd_wearingCity <- distinct(sd_wearingCity)
```

Exclude days where wear = 0 or mvpa > wear (deletes one participant from Saskatoon in wave 1 -- only has one minute of accelerometer data but wear == 0)
```{r}
# 1. all person-days
  # there are 5671 person-days total. We'll filter out the following 106 person-days to yield a TOP with 5565 person-days:
        # (a) more minutes of mvpa than wear time (21 person-days)
              length(sd_all[sd_all$mean_mvpa_sd > sd_all$wear,]$interact_id)
        # (b) the same number of wear and mvpa minutes with at least 1 sedentary or lightly active minutes (85 person-days)  
              length(sd_all[sd_all$mean_mvpa_sd == sd_all$wear & !(sd_all$mean_light_sd == 0 & sd_all$mean_sed_sd == 0),]$interact_id)
    # there are also 96 person-days with 0 minutes of wear time but they're all removed by filtering out the above
      length(sd_all[sd_all$wear==0,]$interact_id) # 96 person-days with 0 wear minutes

  # can replicate the original TOP (5559 person-days) by filtering out those where MVPA > wear (using original activity_levels variable)
    length(sd_all[sd_all$mean_mvpa_sdOld < sd_all$wear,]$interact_id) # 5559
      # difference of 6: 4 with MVPA > wear using old but not new logic, 2 with mvpa == wear but 0 light+sed and thus shouldn't be deleted

  # apply logic using the *new* activity_levels variable (activity_levels_shelby) yields 5565 person-days (described above)
    sd_all<-sd_all[sd_all$mean_mvpa_sd < sd_all$wear | 
              # adding the following logic to keep the 2 person-days where mvpa == wear with 0 sedentary and 0 light active minutes
                (sd_all$mean_mvpa_sd == sd_all$wear & sd_all$mean_light_sd == 0 & sd_all$mean_sed_sd == 0),]

# 2. person-days where wearing == 1
  # apply logic using the *new* activity_levels variable (activity_levels_shelby) to the wearing file (filtered on wearing but not in_city)
    sd_wearing<-sd_wearing[sd_wearing$mean_mvpa_sd < sd_wearing$wear | 
            # adding the following logic to keep the 2 person-days where mvpa == wear with 0 sedentary and 0 light active minutes
              (sd_wearing$mean_mvpa_sd == sd_wearing$wear & sd_wearing$mean_light_sd == 0 & sd_wearing$mean_sed_sd == 0),]

    # fewer people are filtered out (vs. sd_all) because sd_all doesn't filter out wearing == 0 OR in_city == 0
    # mvpa is thus higher in sd_all vs. sd_all test, because the former includes any minutes of MVPA where wearing == 0
    # therefore more times where mvpa > wear

# 3. person-days where wearing == 1 & in_city == 1
  # are there any additional days in the sd_wearingCity data frame that need to be removed? 
  # No, there are 8 person-days where mvpa == wear but there were 0 light and sedentary minutes
      length(sd_wearingCity[sd_wearingCity$mean_mvpa_sd == sd_wearingCity$wear,]$interact_id) # 8 where wear == mvpa 
      length(sd_wearingCity[sd_wearingCity$mean_mvpa_sd == sd_wearingCity$wear & 
                          !(sd_wearingCity$mean_light_sd == 0 & sd_wearingCity$mean_sed_sd == 0),]$interact_id)
      length(sd_wearingCity[sd_wearingCity$mean_mvpa_sd > sd_wearingCity$wear,]$interact_id) # 0 where mvpa > wear
```

*NEW*: compare mean daily MVPA among (a) all person-days and (b) person-days where in_city == 1 and wearing == 1, using (a) activity_levels from the TOP ("old_logic"), and (b) activity_levels_shelby, calculated using code from the INTERACT baseline paper ("new_logic") 
```{r}
# 1. all person-days, not filtered by in_city or wearing
  sd_wave <- sd_all %>% group_by(wave_id) %>%
                        summarize(
                          mvpa_all_old_logic = mean(mean_mvpa_sdOld, na.rm = TRUE), # note these (approximately) match paper 1
                          mvpa_all_new_logic = mean(mean_mvpa_sd, na.rm = TRUE))
  sd_wave<-as.data.frame(sd_wave)

# 2. filtered to wearing == 1
  wearing_wave <- sd_wearing %>% group_by(wave_id) %>%
                                 summarize(mvpa_new_logic_wearing1 = mean(mean_mvpa_sd, na.rm = TRUE))
  wearing_wave<-as.data.frame(wearing_wave)

# 3. filtered to in_city == 1 and wearing == 1
  sd_wearingCity_wave <- sd_wearingCity %>% group_by(wave_id) %>%
                      summarize(mvpa_new_logic_wearing_inCity1 = mean(mean_mvpa_sd, na.rm = TRUE))
                               # note this matches baseline INTERACT paper
  sd_wearingCity_wave<-as.data.frame(sd_wearingCity_wave)

# merge together to compare
  merge(sd_wave,merge(wearing_wave,sd_wearingCity_wave,by=c("wave_id")),by="wave_id")
  rm(sd_wave,sd_wearingCity_wave,wearing_wave)
```

*NEW*: compare mean mvpa by city and wave among (a) all person-days and (b) person-days where in_city == 1 and wearing == 1, using (a) activity_levels from the TOP ("old_logic"), and (b) activity_levels_shelby, calculated using code from the INTERACT baseline paper ("new_logic") 
```{r}
# 1. all person-days, not filtered by in_city or wearing
  sd_table <- sd_all %>% group_by(city_id,wave_id) %>%
                         summarize(mvpa_all_old_logic = mean(mean_mvpa_sdOld, na.rm = TRUE), # do these match paper 1?
                                   mvpa_all_new_logic = mean(mean_mvpa_sd, na.rm = TRUE))
  sd_table<-as.data.frame(sd_table)

# 2. person-days where wearing == 1
  wearing_table <- sd_wearing %>% group_by(city_id,wave_id) %>%
                                  summarize(mvpa_new_logic_wearing1 = mean(mean_mvpa_sd, na.rm = TRUE))
  wearing_table<-as.data.frame(wearing_table)

# 3. person-days where wearing == 1 & in_city == 1
  sd_wearingCity_table <- sd_wearingCity %>% group_by(city_id,wave_id) %>%
                                             summarize(mvpa_new_logic_wearing_inCity1 = mean(mean_mvpa_sd, na.rm = TRUE))
                                                  # note these match the baseline INTERACT paper
  sd_wearingCity_table<-as.data.frame(sd_wearingCity_table)

  merge(sd_table,merge(wearing_table,sd_wearingCity_table,by=c("wave_id","city_id")),by=c("wave_id","city_id"))
  rm(sd_table,sd_wearingCity_table,wearing_table)
```

Generate a flag indicating if a participant contibuted data in both waves
```{r}
# 1. all person days (excluding days as outlined above)
  sd_all<-sd_all %>% group_by(interact_id) %>% mutate(n=length(unique(wave_id))) 
  sd_all$both<-ifelse(sd_all$n==1,"no","yes")
    length(unique(sd_all[sd_all$n==2,]$interact_id))

# 2. person days where wearing == 1
  sd_wearing<-sd_wearing %>% group_by(interact_id) %>% mutate(n=length(unique(wave_id))) 
  sd_wearing$both<-ifelse(sd_wearing$n==1,"no","yes")
    length(unique(sd_wearing[sd_wearing$n==2,]$interact_id))

# 3. person days where wearing == 1 & in_city == 1
  sd_wearingCity<-sd_wearingCity %>% group_by(interact_id) %>% mutate(n=length(unique(wave_id))) 
  sd_wearingCity$both<-ifelse(sd_wearingCity$n==1,"no","yes")
    length(unique(sd_wearingCity[sd_wearingCity$n==2,]$interact_id))
```

## 4. Add health data

### Clean data from health survey

Income:  
```{r}
  health$income<-factor(ifelse(health$income %in% c("1","2","3","4","5","6","7"),"<$50,000",
                        ifelse(health$income==8,"$50,000 to $99,999",
                        ifelse(health$income %in% c(9,10,11),"$100,000 or more",
                        ifelse(health$income==77,"don't know/prefer not to answer",NA)))),
                  levels=c("<$50,000","$50,000 to $99,999","$100,000 or more",
                           "don't know/prefer not to answer"))
  table(health$income)
```

Gender:  
```{r}
  health$gender<-ifelse(health$gender == 1,"man",
                 ifelse(health$gender == 2,"woman","other"))
  table(health$gender)
```

Children: 
```{r}
  health$children<-as.factor(ifelse(health$children==1,"yes","no"))
  summary(as.factor(health$children))
```

Marital status
```{r}
  health$marital_status<-as.factor(ifelse(health$marital_status==1,"single",
                                   ifelse(health$marital_status==2,"married",
                                   ifelse(health$marital_status==3,"separated",
                                   ifelse(health$marital_status==4,"widowed",NA)))))
  summary(as.factor(health$marital_status))
```

Employment status
```{r}
  health$employment<-as.factor(ifelse(health$employment==1,"retired and not working",
                               ifelse(health$employment %in% c(2,3),"employed",
                               ifelse(health$employment %in% c(4,5),"unemployed",
                               ifelse(health$employment==6,"other",NA)))))
  summary(as.factor(health$employment))
```

Education: 
```{r}
  health$education<-as.factor(ifelse(health$education==1,"elementary school",
                              ifelse(health$education==2,"secondary school",
                              ifelse(health$education==3,"trade school or college",
                              ifelse(health$education %in% c(4,5),"university or graduate degree",
                              ifelse(health$education==77,"don't know/prefer not to answer",NA))))))
  summary(as.factor(health$education))
```

Dwelling type: 
```{r}
 health$dwelling_type<-as.factor(ifelse(health$dwelling_type==1,"single-detached",
                                 ifelse(health$dwelling_type==2,"semi-detached",
                                 ifelse(health$dwelling_type==3,"row",
                                 ifelse(health$dwelling_type==4,"apartment duplex",
                                 ifelse(health$dwelling_type==5,"apartment <5 stories",
                                 ifelse(health$dwelling_type==6,"apartment 5+ stories",
                                 ifelse(health$dwelling_type==7,"mobile home",
                                 ifelse(health$dwelling_type==8,"seniors home",
                                 ifelse(health$dwelling_type==9,"other",
                                 ifelse(health$dwelling_type==77,"don't know/prefer not to answer",NA)))))))))))
```

### Merge health and SD data

```{r}
# 1. all person-days
  sd_all<-merge(sd_all,health,by=c("interact_id","wave_id"),all.x=TRUE) # 5565
# 2. person-days where wearing == 1
  sd_wearing<-merge(sd_wearing,health,by=c("interact_id","wave_id"),all.x=TRUE) # 5575
# 3. person-days where wearing == 1 and in_city == 1
  sd_wearingCity<-merge(sd_wearingCity,health,by=c("interact_id","wave_id"),all.x=TRUE) # 5395
  rm(health)
```

Clean wave id (makes it easier for figures)
```{r}
# 1. all person-days
  sd_all$wave_id<-ifelse(sd_all$wave_id==1,"wave 1","wave 2")
# 2. person-days where wearing == 1
  sd_wearing$wave_id<-ifelse(sd_wearing$wave_id==1,"wave 1","wave 2")
# 3. person-days where wearing == 1 & in_city == 1
  sd_wearingCity$wave_id<-ifelse(sd_wearingCity$wave_id==1,"wave 1","wave 2")
```

## 5. Add stringnecy index

Read in data
```{r}
  si<-read.csv("./Data/IRPP stringency index/2021 09 22 INDEX Centre of Excellence COVID Policy Data_New Codebook.csv")
  si$date<-as.Date(si$date)
  si$province<-si$Province.Territory
```

Add province to sd data 
```{r}
# 1. all person days
  sd_all$province<-ifelse(sd_all$city_id=="vancouver","British Columbia",
                   ifelse(sd_all$city_id=="saskatoon","Saskatchewan",
                   ifelse(sd_all$city_id=="montreal","Quebec",NA)))
# 2. person-days where wearing == 1
  sd_wearing$province<-ifelse(sd_wearing$city_id=="vancouver","British Columbia",
                       ifelse(sd_wearing$city_id=="saskatoon","Saskatchewan",
                       ifelse(sd_wearing$city_id=="montreal","Quebec",NA)))
# 3. person-days where wearing == 1 & in_city == 1 
  sd_wearingCity$province<-ifelse(sd_wearingCity$city_id=="vancouver","British Columbia",
                           ifelse(sd_wearingCity$city_id=="saskatoon","Saskatchewan",
                           ifelse(sd_wearingCity$city_id=="montreal","Quebec",NA)))
```

Merge with sd data
```{r}
# 1. all person days
  sd_all<-merge(sd_all,si[si$date %in% sd_all$date,colnames(si) %in%
                          c("date","province","stringencyIndex")],by=c("date","province"),all.x=TRUE)
# 2. person-days where wearing == 1
  sd_wearing<-merge(sd_wearing,si[si$date %in% sd_wearing$date,colnames(si) %in%
                          c("date","province","stringencyIndex")],by=c("date","province"),all.x=TRUE)
# 3. person-days where wearing == 1 & in_city == 1
  sd_wearingCity<-merge(sd_wearingCity,si[si$date %in% sd_wearingCity$date,colnames(si) %in%
                          c("date","province","stringencyIndex")],by=c("date","province"),all.x=TRUE)
  rm(si)
```

## 6. Add weather data

Read in weather data from 2016 to December 31, 2020
```{r}
  weather<-read.csv("Data/weather data/2016_2020_weather_vic_van_sask_mtl.csv")
```

Read in weather data from Jan and Feb 2021 (all cities) and 2020 for BC (to fill in the gaps where these data are missing above)
```{r}
# 2021
  weatherBC2021<-read.csv("Data/weather data/en_climate_daily_BC_1108395_2021_P1D.csv")
    weatherBC2021$city_id<-"Vancouver"
    colnames(weatherBC2021)<-colnames(weather)
  weatherQC2021<-read.csv("Data/weather data/en_climate_daily_QC_702S006_2021_P1D.csv") 
    weatherQC2021$city_id<-"Montreal"
    colnames(weatherQC2021)<-colnames(weather)
  weatherSK2021<-read.csv("Data/weather data/en_climate_daily_SK_4057165_2021_P1D.csv") 
    weatherSK2021$city_id<-"Saskatoon"
    colnames(weatherSK2021)<-colnames(weather)

# 2020
  weatherBC2020<-read.csv("Data/weather data/en_climate_daily_BC_1108395_2020_P1D.csv")
    weatherBC2020$city_id<-"Vancouver"
    colnames(weatherBC2020)<-colnames(weather)
  # remove dates with missing temperature data 
    weatherBC2020<-weatherBC2020[!is.na(weatherBC2020$total_precip_mm) & !is.na(weatherBC2020$mean_temp_c),]
  # filter to the dates in 2020 with missing temperature or precipitation data 
    missing<-weather[(is.na(weather$total_precip_mm) | is.na(weather$mean_temp_c)) & 
                      weather$city_id=="Vancouver" & weather$date > "2020-09-01",]
    weatherBC2020<-weatherBC2020[weatherBC2020$date_time %in% missing$date_time,]
```

Combine old weather data with updated weather data
```{r}
  weather<-rbind(weather,weatherBC2021,weatherQC2021,weatherSK2021)
  weather<-weather[!is.na(weather$total_precip_mm) & !is.na(weather$mean_temp_c),]
  weather<-rbind(weather,weatherBC2020)
  weather<-weather[weather$date_time<"2021-03-01",]
  rm(weatherBC2020,weatherBC2021,weatherQC2021,weatherSK2021,missing)
```

Reformat column names to match the sd_all dataframe (format city_id to be all lowercase, rename datetime to date)
```{r}
  weather$city_id<-tolower(weather$city_id)
  weather<-weather %>% rename(date=date_time,
                              mean_temp=mean_temp_c,
                              max_temp=max_temp_c,
                              min_temp=min_temp_c,
                              total_precip=total_precip_mm)
```
Explore variables available
```{r}
  weather<-weather %>% dplyr::select(date,mean_temp,total_precip,city_id)
```

Merge with sd data
```{r}
# 1. all person-days
  sd_all<-merge(sd_all,weather,by=c("date","city_id"),all.x=TRUE)

# 2. person days where wearing == 1
  sd_wearing<-merge(sd_wearing,weather,by=c("date","city_id"),all.x=TRUE)

# 3. person days where wearing == 1 & in_city == 1
  sd_wearingCity<-merge(sd_wearingCity,weather,by=c("date","city_id"),all.x=TRUE)
  rm(weather)
```

## 7. Export data

```{r}
  write.csv(sd_all,"sensdoc-all-2023.csv",row.names = FALSE)
  write.csv(sd_wearingCity,"sensdoc-wearingCity-2023.csv",row.names = FALSE)
  write.csv(sd_wearing,"sensdoc-wearing-2023.csv",row.names = FALSE)
```

### 2023-10-12 paper1-INTERACT-analysis.Rmd
Update objective 2 to run on valid wear days (sd_valid)

@@
 ```{r,results="hide"}
-  sd_pa<-read.csv("sensdoc-wearing-2023.csv") # 5575
+  sd_pa<-read.csv("Data//final-for-analysis//sensdoc-wearing-2023.csv") # 5575
 ```
@@
 
-  ggsave(si_plot,file="figures//figure4.pdf",width=5.5,height=3.25)
-  ggsave(si_plot,file="submission to IJE//figure4.pdf",width=175,height=95,unit="mm") # for IJBNPA
+  ggsave(si_plot,file="figures//figure4.png",width=5.5,height=3.25)
+  ggsave(si_plot,file="submission to IJE//figure4.png",width=175,height=95,unit="mm") # for IJBNPA
 ```
@@
 
-*Columns 5 and 6: valid participant-days from wave 2, stratified by stringency index (sd_valid)*: 
+*Columns 5, 6 and 7: valid participant-days from wave 2, stratified by stringency index (sd_valid)*: 
 CLEAR imbalances of gender, age, kids, education and employment status (not income)
@@
 
-  ggsave(figure1,file="figures//figure1.pdf",width=6,height=4)
-  ggsave(figure1,file="submission to IJE//figure1.pdf",width=175,height=100,unit="mm")
+  ggsave(figure1,file="figures//figure1.png",width=6,height=4)
+  ggsave(figure1,file="submission to IJE//figure1.png",width=175,height=100,unit="mm")
 ```
@@
   figure2
-  ggsave(figure2,file="figure2.pdf",width=6,height=4)
-  ggsave(figure2,file="submission to IJE//figure2.pdf",width=175,height=120,unit="mm")
+  ggsave(figure2,file="figure2.png",width=6,height=4)
+  ggsave(figure2,file="submission to IJE//figure2.png",width=175,height=120,unit="mm")
 ```
@@
 
-**Create income plot**
 ```{r,echo=FALSE,results=FALSE,warning=FALSE}
+# income plot
+
 # Add a income-wave interaction term to the adjusted model  
@@
 
-**Create education plot**
 ```{r,echo=FALSE,results=FALSE,warning=FALSE}
+# education plot
+
 # Add a education-wave interaction term to the adjusted model 
@@
 
-**Create employment figure**
 ```{r,echo=FALSE,results=FALSE,warning=FALSE}
+# employment plot
+
 # Add an employment status-wave interaction term to the adjusted model 
@@
             nrow=1)
-ggsave(em_figures,file="figures//figure3.jpeg",width=175,height=85,units="mm")
-ggsave(em_figures,file="figures//figure3.pdf",width=175,height=85,units="mm")
-ggsave(em_figures,file="submission to IJE//figure3.pdf",width=175,height=85,units="mm")
+ggsave(em_figures,file="figures//figure3.png",width=175,height=85,units="mm")
+ggsave(em_figures,file="submission to IJE//figure3.png",width=175,height=85,units="mm")
 ```
@@
 #### Unadjusted model (row 1 table 3)
-```{r,echo=FALSE,result=FALSE,warning=FALSE}
-  length(sd_pa[sd_pa$wave_id=="wave 2",]$interact_id) # missing 7 observations???
-  length(sd_pa[is.na(sd_pa$stringencyIndex) & sd_pa$wave_id=="wave 2",]$interact_id)
-```
 
@@
   si_all_unadj<-glmer.nb(mean_mvpa_sd~ stringencyIndex + (1|interact_id),
-                         data=sd_pa[sd_pa$wave_id=="wave 2",]) 
-  summary(si_all_unadj)
+                         data=sd_valid[sd_valid$wave_id=="wave 2",]) 
+  # summary(si_all_unadj)
 ```
@@
                           (1|interact_id),
-                          data=sd_pa[sd_pa$gender!="other" & sd_pa$age_cat!="missing" &
-                                     sd_pa$wave_id=="wave 2",])
-  summary(si_all_cov_fully)
+                          data=sd_valid[sd_valid$gender!="other" & sd_valid$age_cat!="missing" &
+                                     sd_valid$wave_id=="wave 2",])
 ```
@@
 
-*Objective 1*
+**Objective 1**
 
@@
 ```
+
 Adjusted model (complete case; no missing data for confounders)
@@
 
-*Sensitivity analysis 1*
+**Sensitivity analysis 1**
 
@@
 ```{r}
-print("age")
 summary(sd_valid_both$age_cat)
-
-print("mean_temp")
 summary(sd_valid_both$mean_temp)
-
-print("total_precip")
 summary(sd_valid_both$total_precip)
 ```
-*Sensitivity analysis 2*
+
+**Sensitivity analysis 2**
 
@@
 ```{r}
-print("age")
 summary(sens_valid$age_cat) # don't need to check for missingness in weather data since we aren't controlling for weather in this model
@@
 
-*Objective 2*
+**Objective 2**
 
@@
 ```{r}
-print("age")
 summary(sd_valid[sd_valid$wave_id=="wave 2",]$mean_temp)

### 2023-10-12 paper1-INTERACT-data-prep.Rmd
Clean data prep code

@@
 ```
+
 Explore variables available
@@
   sd_wearing<-merge(sd_wearing,weather,by=c("date","city_id"),all.x=TRUE)
-
-# 3. person days where wearing == 1 & in_city == 1
-  sd_wearingCity<-merge(sd_wearingCity,weather,by=c("date","city_id"),all.x=TRUE)
-  rm(weather)
 ```
@@
 ```{r}
-  write.csv(sd_all,"sensdoc-all-2023.csv",row.names = FALSE)
-  write.csv(sd_wearingCity,"sensdoc-wearingCity-2023.csv",row.names = FALSE)
-  write.csv(sd_wearing,"sensdoc-wearing-2023.csv",row.names = FALSE)
+  write.csv(sd_all,"Data//final-for-analysis//sensdoc-all-2023.csv",row.names = FALSE)
+  write.csv(sd_wearing,"Data//final-for-analysis//sensdoc-wearing-2023.csv",row.names = FALSE)
 ```

### 2023-10-12 paper1-INTERACT-analysis.Rmd
Removed code to filter out gender = other (n=16) in adjusted model for objective 2

@@
                           (1|interact_id),
-                          data=sd_valid[sd_valid$gender!="other" & sd_valid$age_cat!="missing" &
-                                     sd_valid$wave_id=="wave 2",])
+                          data=sd_valid[sd_valid$wave_id=="wave 2",])
 ```

### 2024-09-09 paper1-INTERACT-analysis.Rmd
updated figures to read "phase" instead of "wave", to match with new language in the paper. Also reviewed the code to ensure that models were adjusted for the correct variables as I was revising the proofs. I noticed that the fully adjusted model for objective one was missing education level, so I added that in and re-ran the model. It was also missing from the first of the three interaction plots. Adding this in did not meaninfully change the results -- only the CI shifted by 0.01. Similarly, I noticed that income was missing from the code for the fully adjusted model for objective 2. I added

@@
 ```
+total unique particiapnts across both waves
+```{r}
+  length(unique(sd_ppl$interact_id))
+  table(sd_ppl$wave_id)
+```
 
@@
   table(sd_ppl_valid$wave_id) # one participant from Saskatoon in wave 1 has gone missing?
+```
+
+total unique particiapnts across both waves
+```{r}
+  length(unique(sd_ppl_valid$interact_id))
 ```
@@
                  plot.title = element_text(face="bold",size=10,hjust=0.5)) +
-           annotate(geom="text", x=as.Date('2020-12-04'), y=10, label="Wave 2 data collection period",
+           annotate(geom="text", x=as.Date('2020-12-04'), y=10, label="Phase 2 data collection period",
                     color="grey30",size=2.75)
@@
 
-  ggsave(si_plot,file="figures//figure4.png",width=5.5,height=3.25)
-  ggsave(si_plot,file="submission to IJE//figure4.png",width=175,height=95,unit="mm") # for IJBNPA
+  ggsave(si_plot,file="figures//suppFigure3-20240909.png",width=5.5,height=3.25)
 ```
@@
 ```{r,echo=FALSE}
-wave.labs<-c("Wave 1","Wave 2")
+wave.labs<-c("Phase 1","Phase 2")
 names(wave.labs)<-c("wave 1","wave 2")
@@
          y="Frequency",
-         title="Wear time" #,
-         #subtitle="Wave 1 (dark blue) vs. Wave 2 (light blue)"
-         ) +
-    geom_vline(xintercept=600,color="black",linetype="dashed") +
+         title="Wear time") +
+    #geom_vline(xintercept=600,color="black",linetype="dashed") +
     facet_grid(vars(wave_id),
                labeller = labeller(wave_id=wave.labs)) +
-    # scale_fill_manual(values=c("wave 1"="#0a9ffe", # green = "#81DA66"
-    #                             "wave 2"="#fe7701"), # darker orange #fe7701 #ff8c09
-    #                    labels=c("Wave 1","Wave 2")) + 
-    scale_fill_grey(labels=c("Wave 1","Wave 2")) +
-    scale_color_grey(labels=c("Wave 1","Wave 2"))+
+    scale_fill_grey(labels=c("Phase 1","Phase 2")) +
+    scale_color_grey(labels=c("Phase 1","Phase 2"))+
     theme(axis.title.x = element_text(face="bold",
@@
          y="Frequency",
-         title="Moderate-to-vigorous physical activity" #,
-         #subtitle="Wave 1 (dark blue) vs. Wave 2 (light blue)"
-         ) +
+         title="Moderate-to-vigorous physical activity") +
     facet_grid(vars(wave_id),
                labeller = labeller(wave_id=wave.labs)) +
-    # scale_fill_manual(values=c("wave 1"="#0a9ffe", # green = "#81DA66"
-    #                             "wave 2"="#fe7701"), # darker orange #fe7701 #ff8c09
-    #                    labels=c("Wave 1","Wave 2")) + 
-    scale_fill_grey(labels=c("Wave 1","Wave 2")) +
-    scale_color_grey(labels=c("Wave 1","Wave 2"))+
+    scale_fill_grey(labels=c("Phase 1","Phase 2")) +
+    scale_color_grey(labels=c("Phase 1","Phase 2"))+
     theme(axis.title.x = element_text(face="bold",
@@
 
-  # titleMain=ggpubr::text_grob("Distribution of daily minutes of wear time and moderate-to-vigorous physical activity:\nWave 1 (2018-2019) vs. Wave 2 (2020-2021)\n",
-  #                             size =10,face="bold")  
   titleY=ggpubr::text_grob("Frequency",
@@
 
-  figure1<-grid.arrange(wear_plot,mvpa_plot,ncol=2,
+  suppFigure2<-grid.arrange(wear_plot,mvpa_plot,ncol=2,
                #top=titleMain, # removing title for IJBNPA (titles should be in the text, not in the figure)
@@
 
-  ggsave(figure1,file="figures//figure1.png",width=6,height=4)
-  ggsave(figure1,file="submission to IJE//figure1.png",width=175,height=100,unit="mm")
-```
-
-**Figure 2:** Trend in daily minutes of moderate-to-vigorous physical activity for each wave
+  ggsave(suppFigure2,file="figures//suppFigure2-20240909.png",width=6,height=4)
+```
+
+**Figure 1:** Trend in daily minutes of moderate-to-vigorous physical activity for each wave
 ```{r,echo=FALSE,warning=FALSE}
@@
 
-  figure2<-ggplot(sd_valid[sd_valid$wave_id=="wave 1",],aes(x=doy3,y=mean_mvpa_sd,color=wave_id)) +
+  figure1<-ggplot(sd_valid[sd_valid$wave_id=="wave 1",],aes(x=doy3,y=mean_mvpa_sd,color=wave_id)) +
                           geom_point(alpha=0.15,size=1)+
@@
                                                        "wave 2"="black"),
-                                              labels=c("Wave 1 (2018-2019)","Wave 2 (2020-2021)")) +
+                                              labels=c("  Phase 1 (2018-2019)","  Phase 2 (2020-2021)")) +
                           geom_smooth(method=lm)+
@@
                           theme_classic()+
-                          labs(#title = "Daily minutes of moderate-to-vigorous physical activity by day of the year: ", # removing title for IJBNPA
-                               #subtitle = "Wave 1 (2018-2019) vs. Wave 2 (2020-2021)",
-                               y = "Daily minutes of moderate-to-vigorous physical activity",
+                          labs(y = "Daily minutes of moderate-to-vigorous physical activity",
                                x = "Date")+
@@
                                 legend.text = element_text(size=8,margin = margin(r = 0.4, unit = 'cm'))) 
-  figure2
-  ggsave(figure2,file="figure2.png",width=6,height=4)
-  ggsave(figure2,file="submission to IJE//figure2.png",width=175,height=120,unit="mm")
+  figure1
+  ggsave(figure1,file="figures//figure1-20240909.png",width=6,height=4.5)
 ```
@@
 
-*Model: Adjusting for city, age, employment status, income, temp, precipitation and wear time*
+*Model: Adjusting for city, age, employment status, income, education, temp, precipitation and wear time*
 ```{r,warning=FALSE}
@@
                                                  age_cat + employment + income + 
-                                                 mean_temp + total_precip + 
+                                                 mean_temp + total_precip + education +
                                                  (1|interact_id),
                           data=sd_valid[sd_valid$age_cat!="missing",])
+summary(wave_covar_valid_fully)
 ```
@@
                           data=sd_valid_both) 
+summary(both_valid_adj_wear)
 ```
@@
 
-#### Plot effect modifiers of interest (Figure 3): Income, employment status and education
+#### Plot effect modifiers of interest (Figure 2): Income, employment status and education
 
@@
 income_covar_valid_final<-glmer.nb(mean_mvpa_sd~ wave_id + city_id + wave_id*income + wearRS +
-                                                age_cat + employment + income +
+                                                age_cat + employment + income + education +
                                                 mean_temp + total_precip + 
@@
               levels=c("wave 1","wave 2"),
-              labels=c("Wave 1", "Wave 2"))
+              labels=c("Phase 1", "Phase 2"))
 
@@
 
-**Create Figure 3**
+**Create Figure 2**
 ```{r,echo=FALSE,warning=FALSE}
@@
             nrow=1)
-ggsave(em_figures,file="figures//figure3.png",width=175,height=85,units="mm")
-ggsave(em_figures,file="submission to IJE//figure3.png",width=175,height=85,units="mm")
+ggsave(em_figures,file="figures//figure2-20240909.png",width=175,height=85,units="mm")
 ```
@@
 ```{r,warning=FALSE}
-  si_all_cov_fully<-glmer.nb(mean_mvpa_sd~ stringencyIndex + wear + city_id + 
+  si_all_cov_fully<-glmer.nb(mean_mvpa_sd~ stringencyIndex + city_id + 
                           mean_temp + total_precip + 
@@
                           children + 
-                          employment + 
+                          employment + income +
                           (1|interact_id),

### 2024-09-16 paper1-INTERACT-analysis.Rmd
Added additional calculations

@@
   #print(kableone(balance_ppl_by_wave_full,quote=TRUE,noSpaces=TRUE))
+```
+
+Valid days of data per person -- wave 1
+```{r}
+sd_valid_per_person<-sd_valid %>% group_by(interact_id,wave_id) %>% mutate(rows=length(interact_id)) %>% 
+                     dplyr::select(interact_id,rows) %>% distinct()
+mean(sd_valid_per_person[sd_valid_per_person$wave_id=="wave 1",]$rows)
+sd(sd_valid_per_person[sd_valid_per_person$wave_id=="wave 1",]$rows)
+```
+
+Valid days of data per person -- wave 2
+```{r}
+mean(sd_valid_per_person[sd_valid_per_person$wave_id=="wave 2",]$rows)
+sd(sd_valid_per_person[sd_valid_per_person$wave_id=="wave 2",]$rows)
 ```
